---
title: "KEGG_Pathway_Analysis"
author: "HM Putnam"
date: "9/20/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## KEGG enrichment analysis
## Obtained KEGG ontology terms by using KofamScan in command-line; see step 3 in E. Chille lab notebook https://github.com/echille/E.-Chille-Open-Lab-Notebook/blob/master/_posts/2020-10-08-M-capitata-functional-annotation-pipeline.md for instructions. 
## If run into issues with downloading KofamScan software, etc.; see https://github.com/Putnam-Lab/Lab_Management/issues/21

# INFORMATION FOR KEGG ENRICHMENT ANALYSIS FOR R FOUND HERE:
# https://bioinformatics-core-shared-training.github.io/cruk-summer-school-2018/RNASeq2018/html/06_Gene_set_testing.nb.html 
#http://yulab-smu.top/clusterProfiler-book/chapter6.html#kegg-over-representation-test


```{r}
# LOAD PACKAGES

#need to download clusterProfiler package from devtools because it wasn't compatible with new rstudio version
#devtools::install_github(c("GuangchuangYu/DOSE", "GuangchuangYu/clusterProfiler"))
library(clusterProfiler)
library(KEGGprofile) # BiocManager::install("KEGGprofile")
library(KEGGREST)
library(tidyr)
library(stringr)
library(forcats)
library(ggplot2)
library(scales)
library(ape)
library(data.table)
library(tidyverse)
library(fBasics)
```
```{r}

# LOAD DATA

# Pdam_genome_dataframe - what is this? after KEGG enrichment analysis this script with unlist genes within enriched pathways and merge IDs to obtain the gene names 
#to search KEGG organism list, go to: https://www.genome.jp/kegg/catalog/org_list.html
#Hollie, I choose Pocillopora damicornis from the Cnidarians section on the genome list, think it is closest we will get
#Pocillopora damicornis is the closest Cnidarian reference in the kegg enrichment database

Pdam_genome <- keggList("pdam") # call the P. damicornis genome! - notice the pdam terms are rownames!
Pdam_genome_dataframe <- as.data.frame(Pdam_genome) %>%  rownames_to_column() # with will allow us to merge 
colnames(Pdam_genome_dataframe) <- c('gene_name') # rename the columns

# load the Pver annotation files to use downstream in KEGG pathway analysis 
Pver_annot <- read_tsv( "RAnalysis/Genome/Pver_gene_annotation_file.txt", col_names = TRUE) #biological annotation information
colnames(Pver_annot)[2] <- "gene_id" #set column name for consistency across files

#load GFF3
GFF3 <- read.csv("RAnalysis/Genome/Pver_genome_assembly_v1.0.gene.gff3", header = FALSE, sep="", skip = 1)
colnames(GFF3) <- c("prot", "Predict", "id", "start","stop", "pos1", "pos2","pos3", "gene_id") #name columns
GFF3$gene_id <-  gsub(";.*","",GFF3$gene_id) #remove extra characters after the first ;
#There also appears to be the presence of the characters ID= in each row
GFF3$gene_id <- gsub("ID=","",GFF3$gene_id) #remove ID=
#There is also an extra "-g#" that can be removed
GFF3$gene_id <-  gsub("-.*","",GFF3$gene_id) #remove the extra g number after the dash

#crgKEGG_Pgenref_DIAMOND <- read.table(file ="C:/Users/samjg/Documents/Github_repositories/Pgenerosa_TagSeq_Metabolomics/TagSeq/HPC_work/Output/crgKEGG_diamond_out.txt", sep = '\t', header = F)
#colnames(crgKEGG_Pgenref_DIAMOND) <- c('qseqid','sseqid','pident', 'length', 'mismatch', 'gapopen', 'qstart', 'qend', 'sstart', 'send', 'evalue', 'bitscore') # change the names of the columns if the best blast hits

```

```{r}
#need list of FDR and log fold change for genes here, that is in the DESeq2 DEG results list
#loading data for enrichment here
#DEG significant results
DEG.sig <- read.csv("RAnalysis/Output/RNA-seq/DEG/DEGSeq2.sig.results.csv")[,-1]
#how many significant DEGs
nrow(DEG.sig)
#add _gene_ to the gene_id name to match other nomenclature
DEG.sig$gene_id <-  gsub("_g","_gene_g",DEG.sig$gene_id) #add gene to gene_id to match nomenclature 
```

```{r}
# DESEq2 results - assign up and downregualted genes!
#ask Hollie if I want to load all DEGs here or just significant ones

#make new column and dataframe with upregualted or downregulated genes
DEG.sig$DE_dir[DEG.sig$log2FoldChange > 1] <- "up"
DEG.sig$DE_dir[DEG.sig$log2FoldChange < 1]<- "down"
```


# Load KO -- gene mapping info
```{r}

#Load in and prep KofamScan search results
KFS.KO.orig <- read_tsv("RAnalysis/Genome/KEGG-ontology/Pver_KO_annot.tsv.gz", col_names = TRUE)
KFS.KO <- KFS.KO.orig[-1,-1] #remove the first row and column becasue they  contain no data
KFS.KO <- KFS.KO[, c(1,2,5)] #subset the gene ID, KO term, and evalue
colnames(KFS.KO) <- c("gene_id", "ko", "eval") #name columns
#NEED TO DECIDE KO TERM CUTOFF AND WHY OR WHY NOT TO KEEP MORE THAN ONE PER GENE
#FILTER FOR EVALUE AS NEEDED
#KFS.KO <- KFS.KO[,c(1:2)] #remove evalue from the dataframe
Kegg.name <- as.data.frame(unique(KFS.KO$gene_id)) #identify all of the unique gene names
colnames(Kegg.name) <- c("gene_id") #set column name for consistency across files

#Annot appears to have an extra Pver_g1.t2 and are missing the word gene
KFS.KO$gene_id <-  gsub("\\..*","", KFS.KO$gene_id)
KFS.KO$gene_id <-  gsub("Pver_novel_model_","Pver_novel_gene_",KFS.KO$gene_id)
#g# is duplicated after dash 
KFS.KO$gene_id <-  gsub("-.*","",KFS.KO$gene_id) #remove the extra g number after the dash
KFS.KO$gene_id <-  gsub("_gene_g","_g",KFS.KO$gene_id)
KFS.KO$gene_id <-  gsub("_gene_","_",KFS.KO$gene_id)

Kegg.name <- as.data.frame(unique(KFS.KO$gene_id)) #identify all of the unique gene names
colnames(Kegg.name) <- c("gene_id") #set column name for consistency across files

KFS.KO$gene_id <- gsub("Pver_split_", "Pver_", KFS.KO$gene_id)
KFS.KO$gene_id <- gsub("Pver_", "Pver_gene_", KFS.KO$gene_id)
head(KFS.KO)
tail(KFS.KO)
```


```{r}

#clusterProfiler (Yu et al. 2012). clusterprofiler supports direct online access of the current KEGG database, rather than relying on R annotation packages, it also provides some nice visualisation options.

search_kegg_organism('pdam', by='kegg_code')

#need list of FDR and log fold change for genes here, that is in the DESeq2 DEG results list
#loading data for enrichment here
#DEG significant results
DEG.sig <- read.csv("RAnalysis/Output/RNA-seq/DEG/DEGSeq2.sig.results.csv")[,-1]
#how many significant DEGs
nrow(DEG.sig)
#add _gene_ to the gene_id name to match other nomenclature
DEG.sig$gene_id <-  gsub("_g","_gene_g",DEG.sig$gene_id) #remove extra characters

#rename columns for FDR and logFC
#colnames(DEG.sig)[2] <- c("logFC") #rename columns
#colnames(DEG.sig)[6] <- c("FDR") #rename columns
#head(DEG.sig)

# #filter significant DEGs for FDR and logFC cutoffs
# sigGenes <- DEG.sig$gene_id[ DEG.sig$FDR < 0.01 & 
#                               !is.na(DEG.sig$FDR) &
#                               abs(DEG.sig$logFC) > 1 ]
# sigGenes <- na.exclude(sigGenes)
# kk <- enrichKEGG(gene = sigGenes, organism = 'pdam')
# head(kk, n=10)

KEGG_vectore_Pver <- as.vector(DEG.sig)
KEGG_pver <- enrichKEGG(gene = DEG.sig, 
                            organism  = 'pdam', # 'pdam' is Pocillopora damicornis
                            pvalueCutoff = 0.05) 
```


# Perform KEGG enrichment with GOSeq package
```{r}
#Perform goseq
KOwall <- goseq(pwf, GOref$gene_id, gene2cat=GOKO.terms, test.cats=c("KEGG"), method="Wallenius", use_genes_without_cat=TRUE)

```

# Replace NAs with "KEGG" and extract KO terms from results
```{r}
#Cluster1
KO.05<-KOwall$category[KOwall$over_represented_pvalue<.05]
KO.05<-data.frame(KO.05)
colnames(KO.05) <- c("category")
KO.05 <- merge(KO.05, KOwall.C1, by="category")
KO.05$ontology <- replace_na(KO.05$ontology, "KEGG")
KO.05 <- filter(KO.05, ontology=="KEGG")
KO.05 <- KO.05[order(KO.05$ontology, KO.05$over_represented_pvalue, -KO.05$numDEInCat),]
KO.05$term <- as.factor(KO.05$term)
nrow(KO.05)

```

# Add KO definitions
```{r}
#Prep definition data
KFS.KO.def <- subset(KFS.KO.orig, select=c("#","KO", "KO definition"))
colnames(KFS.KO.def) <- c("sig","category", "term")
KFS.KO.def <- filter(KFS.KO.def, sig=="*")
KFS.KO.def <- KFS.KO.def[,c(2:3)]
#Merge with KEGG outp
KO.05 <- unique(left_join(C1.KO.05[,-6], KFS.KO.def, by=c("category")))

```

# Write output KEGG enrichment files
```{r}
write.csv(C1.KO.05, file = "RAnalysis/Output/RNA-seq/GOSeq/KO.05.csv")

```
