---
title: "DEG_DMG_pathway_analysis"
author: "daniellembecker"
date: "2023-01-20"
output: html_document
---


# Molecular Underpinnings Chronic Nutrient Enrichment Project

## Analysis of comparisons between gene ontology enrichment from differentially methylated genes and differentially expressed genes.
## Previous steps for gene expression include RNAseq workflow in Bioinformatics>RNAseq>RNAseq workflow and Differential Gene Expression statistical analysis in RAnalysis>Scripts>RNAseq_Differential_Gene_Expression to make DEG statistical data sheet
## Previous steps for DNA methylation include WGBS workflow in Bioinformatics>WGBS>WGBS workflow and Differential Methylated Genes statistical analysis in RAnalysis>Scripts>WGBS_WGBS.rmd to make DMG statistical dataframe

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Load libraries

library("tidyverse")
library("RColorBrewer")
library("gridExtra")
library("ComplexHeatmap")
library("dplyr")
library("clusterProfiler")
library("pheatmap")
library("magrittr")
library("rtracklayer")
library("GenomicRanges")
library("GSEABase")
library("GOSim")
library("VennDiagram")
library("stats")
library("GO.db")
library("rrvgo")
library("GOSemSim")
library("AnnotationDbi")
```

# Import the data files 
```{r}
#upload go enrichment summary for DEGs
DEG <- read.csv("RAnalysis/Output/RNA-seq/GOSeq/GO.enrichment.summary.table.DEG.csv", header = TRUE, row.names = 1)

#rename category to GOterm
colnames(DEG)[3] ="GOterm" #rename columns

#write csv with bh_adjust
write.csv(DEG, file = "RAnalysis/Output/RNA-seq/GOSeq/GO.DEG.parent.terms.csv")

#upload go enrichment summary for DMGs
DMG <- read.csv("RAnalysis/Output/WGBS/GO.enrichment.summary.table.ontology.10x.csv", header = TRUE, row.names = 1)

#rename category to GOterm
colnames(DMG)[3] ="GOterm" #rename columns

#add column to specific type of measure
DMG <- DMG %>% add_column(type = "DMG.GO.terms")

#adjust p-values 
DMG$bh_adjust <-  p.adjust(DMG$over_represented_pvalue, method="BH") 

#write csv with bh_adjust
write.csv(DMG, file = "RAnalysis/Output/WGBS/GO.DMG.parent.terms.csv")

```

#make Venn Diagram of overlapping GOslim terms for DMG and DEG comparison

```{r}
venn.plot <- venn.diagram(
  x = list(DEG = DEG$term, DMG = DMG$term),
  filename = 'RAnalysis/Output/meth_express_compare/venn_diagram.png',
  category.names = c('DEG GO terms', 'DMG GO terms'),
  output = TRUE,
  col = "black",
	fill = c("dodgerblue", "goldenrod1"),
  cat.default.pos = "outer",
  cat.fontface = "bold",
  cat.pos = c(-18, 18),
          cat.dist = c(0.055, 0.055)
)
```

#subset data sets to see alike terms
```{r}
#compare these datasets, simply pass them to the comparedf() function
#comparedf(DMG, DEG)

#Use summary() to get a more detailed summary

#summary(comparedf(DEG, DMG, by = "term"))

#these examples show unshared characteristics but don't list the actual shared terms
```

```{r}
#use dplyr inner_join instead

common <- intersect(DMG$term, DEG$term) 

common <- as.data.frame(common)

#found three common terms for DMG and DEG

#leukocyte migration involved in inflammatory response (parent term immune system response)

#positive regulation of transcription from RNA polymerase II promoter in response to heat stress (DNA-templated transcription; response to stress)

#fatty acid derivative biosynthetic process (lipid metabolic process)

#main question is why 10x terms have three overlapping but 5x does not
```

#next steps, fill in remaining parent terms for the NAs in these data sets and see the overlap there
#filter datas sets with NAs for their slim terms
```{r}
DMG.NA <- DMG %>%
  filter(is.na(slim_term))

DEG.NA <- DEG %>%
  filter(is.na(slim_term))
```


#Used QuickGO to search and fill in the remaining NA parent terms https://www.ebi.ac.uk/QuickGO/term/GO:1901363
#DMG go plot
```{r}
#upload complete parent term dataframe DMGs and make figures

#upload complete go enrichment summary for DMGs
DMG.complete <- read.csv("RAnalysis/Data/WGBS/DMG_GO_parent_terms_NAs_completed.csv", header = TRUE)

#adjust p-values 
DMG.complete$bh_adjust <-  p.adjust(DMG.complete$over_represented_pvalue, method="BH") 

GO.plot.DMG.complete <-  ggplot(DMG.complete, aes(x = ontology, y = term)) + 
    geom_point(aes(size=numInCat, fill=bh_adjust, colour=bh_adjust)) + 
    scale_size(name="Number of Genes", range=c(5,12), limits=c(1,550))+
    scale_fill_gradient(name = "Adjusted p-value", low="black", high="gray")+
    scale_colour_gradient(name = "Adjusted p-value", low="black", high="gray")+
    xlab("Ontology Category")+
    facet_grid(slim_term ~ ., scales = "free", space='free', labeller = label_wrap_gen(width = 55, multi_line = TRUE))+
    ggtitle("")+
    theme_bw() + 
    theme(panel.border = element_blank(), 
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"),
          strip.background =element_rect(fill="white"),
          strip.text.y = element_text(angle=0, size = 18, face="bold"),
          strip.text.x = element_text(size = 16),
          axis.text.x = element_text(size = 20, angle=45, hjust=1),
          axis.text.y = element_text(size = 16),
          axis.title.x = element_text(size = 22, face="bold"),
          legend.text=element_text(size=16),
          title=element_text(size=22, face="bold"),
          plot.title = element_text(hjust = 0),
          axis.title.y = element_blank())

 
ggsave("RAnalysis/Output/Final_Figures/DMG.complete.slimterms.pdf", GO.plot.DMG.complete , width = 20, height = 49, units = c("in"))


```

































