---
title: "METH_GE_COMPARE"
author: "Danielle M. Becker"
date: "20220720"
output: html_document
editor_options: 
  chunk_output_type: console
---

#### Gene Expression and DNA Methylation comparison ####
## Used to compare general relationship between DNA methylation and gene expression
## Code developed from Javier A Rodriguez-Casariego paper: https://onlinelibrary.wiley.com/doi/pdf/10.1111/mec.16246?casa_token=DGoNlwKZByAAAAAA:i6JDcylh98txNEMQnr6dUTFxKd1sX7xotbSMZJq-DDOyqZGUPHObc30PhaWOAqpEufHMcJJ_wlQbQvbG
## https://github.com/jarcasariego/ACER_clonal_divergence/blob/main/GE_GbM_association/rmd/GE_DNAm_control.Rmd

```{r load packages}

## load packages
library(ComplexHeatmap)
library(clusterProfiler)
library(pheatmap)
library(magrittr)
library(rtracklayer)
library(GenomicRanges)
library(png)
library(GSEABase)
library(GOSim)
library(stats)
library(GO.db)
library(rrvgo)
library(AnnotationDbi)
library(janitor)
library(rtracklayer) # on bioconducter
library(dplyr)
library(edgeR)
library(factoextra)
library(cowplot)
library(adegenet)
library(ggplot2)
library(matrixStats)
library(grid)
library(gridExtra)
library(RColorBrewer)
library(patchwork)
library(stringr)
library(tidyverse)
library(ggpubr)
pal <- brewer.pal(n = 3, name = "PRGn")
```

#### Functions ####
```{r}
# Function for subset 'mat' for the target expression,DNA methylation, and gene attributes modified from @adowneywall
subMat_control <- function(x){
  y <- data.frame(gene_length=log2(x$gene_length),
                  exon=log2(x$exon_count),totalCpG=log2(x$all_count),
                  Methylation=x$mean_Cc,Methylation_CV=x$cv_Cc,
                  Gene_Expression=x$gene_mean_Cc,Gene_Expression_CV=x$gene_cv_Cc)
  return(y)
}
get_density <- function(x, y, ...) {
  dens <- MASS::kde2d(x, y, ...)
  ix <- findInterval(x, dens$x)
  iy <- findInterval(y, dens$y)
  ii <- cbind(ix, iy)
  return(dens$z[ii])
}
perCentCpGCoverage <- function(x){sum(c(mat$cov5_count/mat$all_count)*100 > x)}
# Function for generating three panel contribution plot for the different PCs
suppContriFig <- function(x){
  temp <- list()
  for(i in 1:3){
    temp[[i]] <- fviz_contrib(x, choice = "var", axes = i, top = 10,color = "black",fill = "grey",
                              sort.val = "none")
  }
  return(temp)
}
```

Load metadata
```{r}
#### Data ####
# Read in metadata
meta.data <- read.csv("RAnalysis/Data/RNA-seq/P.verrucosa/metadata.RNAseq.csv", header=T, sep=",", na.string="NA", stringsAsFactors = F) #read in info
colnames(meta.data) <- c("colonyID","treatment","block","Sample_ID")
ID <- meta.data$Sample_ID
```

Processing expression data
```{r}
counts <- read.csv("RAnalysis/Data/RNA-seq/P.verrucosa/Host/Poc_gene_count_matrix.csv", header = TRUE, row.names = 1) #read in genes count matrix
colnames(counts) <- str_sub(colnames(counts), 1, 3) #remove everything after the first _ to just have the sample names
# Order columns by sample name
colnames(counts) <- c("C17","C18","C19","C20","C21","C22","C23","C24","C25","C26","C27","C28","C29","C30","C31","C32","E10","E11","E12","E13","E14","E15","E16","E1","E2","E3", "E4", "E5", "E6", "E7", "E8", "E9")
counts <- counts[, colnames(counts) %in% ID] #intersect with ID to maintain only samples a
counts <- counts[, order(colnames(counts))]
dge_gene <- DGEList(counts,genes = rownames(counts))
test <- as.data.frame(dge_gene)
saveRDS(dge_gene, "RAnalysis/Data/GE_meth_compare/RNA_gene_preNormalization_DGEListObj.Rdata")
```


```{r}
# Create Methylation Summary Table
#SAMPLES E2, E16, AND C19 WERE EXCLUDED DUE TO DEDUPLICATION AND LOW COUNTS
# Read in gene methylation data
gene_dnam <- readRDS("RAnalysis/Output/WGBS/sample_all_methylated_data_10x.Rdata") #read in info

#rename sample.ID to include corresponding E and C in front for enriched and control for later analysis
gene_dnam <- gene_dnam %>%
    mutate(Sample.ID = recode(Sample.ID, '1' = 'E1', '3' = 'E3', '4' = 'E4', '5' = 'E5', '6' = 'E6', '7' = 'E7', '8' = 'E8', '9' = 'E9','10' = 'E10', '11' = 'E11', '12' = 'E12','13' = 'E13', '14' = 'E14', '15' = 'E15', '16' = 'E16', '17' = 'C17', '18' = 'C18', '20' = 'C20', '21' = 'C21', '22' = 'C22', '23' = 'C23', '24' = 'C24', '25' = 'C25', '26' = 'C26', '27' = 'C27', '28' = 'C28', '29' = 'C29', '30' = 'C30', '31' = 'C31', '32' = 'C32'))

#check to see all where renamed correctly
unique(gene_dnam$Sample.ID)

#change from long to wide format
meth <- data.frame (gene_dnam %>%
                   group_by(Sample.ID, gene) %>%
                   summarise(mean = mean(per.meth), n = n()) %>%
                   spread(., key = Sample.ID, value = mean))

#filter for a min 5 Cpg per gene
meth <- meth[which(meth$n > 5),]
cpg10xgene <- meth[,1:2]  
colnames(cpg10xgene) <- c("gene_id", "cov10_count")
rownames(meth) <- meth$gene
meth <- meth[,-1]
meth <- as.matrix(meth)
meth <- meth/100
saveRDS(meth, "RAnalysis/Data/GE_meth_compare/meth_dataset.RData")

meth.test <- data_frame(meth)

## Create a summary of methylation for each gene ##
mean_meth <- rowMeans(meth)

meta.data <- na.omit(meta.data)

# mean gene meth enriched 
subset_meth_E <- meth[, colnames(meth)%in%list_E] #subset data from main m matrix for just enriched data

#test <- subset_meth_E %>% summarise_at(vars(E1:E9), mean, na.rm = TRUE)

mean_meth_E <- rowMeans(subset_meth_E)

# Mean gene meth control
subset_meth_C <- meth[, colnames(meth)%in%list_C] #subset data from main m matrix for just control data

mean_meth_C <- rowMeans(subset_meth_C)

# Coefficient of Variation for enriched samples
sd_meth_E <- rowSds(subset_meth_E)
cv_meth_E <- sd_meth_E/mean_meth_E
 
# Coefficient of Variation for control
sd_meth_C <- rowSds(subset_meth_C)
cv_meth_C <- sd_meth_C/mean_meth_C


# Coefficient of Variation among treatments
meanAmongTrt_meth <- cbind(mean_meth_E,mean_meth_C)
sd_meanAmongTrt_meth <- rowSds(meanAmongTrt_meth)
mean_meanAmongTrt_meth <- rowMeans(meanAmongTrt_meth)
cv_mean_AmongTrt_meth <- sd_meanAmongTrt_meth/mean_meanAmongTrt_meth

# Difference in gene methylation between treatments
diff_Trt_meth <- mean_meth_E-mean_meth_C

# Creat matrix of all CpG summary stats
meth_summary <- cbind(mean_meth,mean_meth_E,mean_meth_C,cv_meth_E,cv_meth_C,cv_mean_AmongTrt_meth,
                      diff_Trt_meth) # bind all data frames with summary stats
meth_summary[is.na(meth_summary)]<-0 #filter for all methylation data over 0
meth_summary <- data.frame(gene_id = rownames(meth_summary),meth_summary) #change gene ID column name
meth_annot <- left_join(cpg10xgene, meth_summary) #join Cpg and meth summary

# Saving full meth_annot data.frame
saveRDS(meth_annot,"RAnalysis/Data/GE_meth_compare/meth_annot_full.RData")

#### Summarize the gene expression data ####
expression <- dge_gene$counts
subset_expression <- expression[, colnames(expression)%in%list_E] 
subset_expression <- subset_expression[rowSums(subset_expression[])>0,]
expression <- expression[rownames(expression) %in% rownames(subset_expression),]
expression_cpm_log <- log2(cpm(expression)+ 1) # log2 cpm transformation
expression <- expression_cpm_log

# Mean gene expression all samples
mean_express <- rowMeans(expression)

# mean gene expression enriched 
subset_express_E <- expression[, colnames(expression)%in%list_E] #subset data from main m matrix for just enriched data

mean_express_E <- rowMeans(subset_express_E) #take mean per gene ID

# Mean gene expression control
subset_express_C <- expression[, colnames(expression)%in%list_C] #subset data from main m matrix for just control data

mean_express_C <- rowMeans(subset_express_C) #take mean per gene ID

# Coefficient of Variation for control gene expression
sd_express_C <- rowSds(subset_express_C)
cv_express_C <- sd_express_C/mean_express_C

# Coefficient of Variation for enriched gene expression
sd_express_E <- rowSds(subset_express_E)
cv_express_E <- sd_express_E/mean_express_E

# Standard Error for control gene expression
se_express_C <- sd_express_C/sqrt(length((subset_express_C)))

# Standard Error for enriched gene expression
se_express_E <- sd_express_E/sqrt(length((subset_express_E)))

# Coefficient of Variation among treatments
meanAmongTrt_express <- cbind(mean_express_E,mean_express_C)
sd_meanAmongTrt_express <- rowSds(meanAmongTrt_express)
mean_meanAmongTrt_express <- rowMeans(meanAmongTrt_express)
cv_mean_AmongTrt_express <- sd_meanAmongTrt_express/mean_meanAmongTrt_express

# Difference in gene expression between Trt
diff_Trt_express <- mean_express_E-mean_express_C

# Create matrix of all gene expression summary stats
gene_summary <- cbind(mean_express,mean_express_E,mean_express_C, se_express_C, se_express_E, cv_express_E,cv_express_C,cv_mean_AmongTrt_express,
                      diff_Trt_express)
gene_summary[is.na(gene_summary)]<-0
colnames(gene_summary) <- paste0("gene_",colnames(gene_summary))
gene_summary <- data.frame(gene_id = rownames(gene_summary),gene_summary)

# transcriptomic data
transcriptomic <- read.csv("RAnalysis/Output/RNA-seq/GOSeq/Host/rename_structure_gff3.csv") #load annotation GFF3 file

#remove first three columns in data frame
transcriptomic <- transcriptomic[,-(1:3),drop=FALSE]
transcriptomic <- transcriptomic[,-(4:6),drop=FALSE]

#add gene between Pver_g to match gene_summary
transcriptomic$gene_id <-  gsub("Pver_g","Pver_gene_g", transcriptomic$gene_id)

#combine gene summmary and gene IDs from annot
gene_annot <- left_join(gene_summary, transcriptomic)

#delete length and gene information that is unnecessary
gene_annot <- gene_annot[,-(9:13),drop=FALSE]

# Saving full gene_annot data.frame
saveRDS(gene_annot,"RAnalysis/Data/GE_meth_compare/gene_annot_full.RData")

# Combine gene expression, exon count and CpG (DNA mehtylation data)
gSum <- left_join(meth_annot, gene_annot)

# Reducing the data.frame to include places were we have all the data.
gSum_red<-gSum[!is.na(gSum$cov10_count),]
gSum_red<-gSum_red[!is.na(gSum_red$gene_mean_express),]

# Saving reduced data.frame
saveRDS(gSum,"RAnalysis/Data/GE_meth_compare/Multi_geneSummaryComplete.RData")

# Saving the reduced data.frame
saveRDS(gSum_red,"RAnalysis/Data/GE_meth_compare/Multi_geneSummaryReduced.RData")
```

#compare global mean meth vs express overall
```{r}
mat <- readRDS("RAnalysis/Data/GE_meth_compare/Multi_geneSummaryReduced.RData")

## Mean Methylation vs Gene Expression following Liew et al
df <- mat

#log gene expression x % methylation
df$log_gene <- log(df$gene_mean_express+1)
df$per_meth <- df$mean_meth * 100

#plot of global express x meth
meth.express.compare <- ggplot(df, aes(x=per_meth, y=log_gene,)) + 
    geom_point() + ylim(0,2) + 
    geom_smooth(method=lm,colour="orange",size=2,na.rm = TRUE) +
    scale_color_viridis() +
    labs(x="DNA Methylation (%)", y = "log10 (mean expression +1)") +
  stat_cor(method = "kendall", digits = 3, label.x = 60, label.y = 0.5 ) +
    theme_cowplot();meth.express.compare

### Figure Mean Methylation vs Gene Expression###
ggsave(meth.express.compare, filename="RAnalysis/Output/Final_Figures/Host/global_express_meth_correlation.png", height = 7, width = 10)

```

## Analysis of comparisons between gene ontology enrichment from differentially methylated genes and differentially expressed genes.

# Import the data files 
```{r}
#upload go enrichment summary for DEGs
DEG <- read.csv("RAnalysis/Data/RNA-seq/P.verrucosa/Host/DEG_GO_parent_terms_complete.csv", header = TRUE)

#adjust p-values 
DEG$bh_adjust <-  p.adjust(DEG$over_represented_pvalue, method="BH") 

#rename category to GOterm
colnames(DEG)[3] ="GOterm" #rename columns

#write csv with bh_adjust
write.csv(DEG, file = "RAnalysis/Output/RNA-seq/GOSeq/Host/GO.DEG.parent.terms.host.csv")

#upload go enrichment summary for DMGs
DMG <- read.csv("RAnalysis/Output/WGBS/GO.enrichment.summary.table.ontology.10x.csv", header = TRUE, row.names = 1)

#rename category to GOterm
colnames(DMG)[3] ="GOterm" #rename columns

#add column to specific type of measure
DMG <- DMG %>% add_column(type = "DMG.GO.terms")

#adjust p-values 
DMG$bh_adjust <-  p.adjust(DMG$over_represented_pvalue, method="BH") 

#write csv with bh_adjust
write.csv(DMG, file = "RAnalysis/Output/WGBS/GO.DMG.parent.terms.csv")

```


#Filled in all NAs with parent terms for completed data frames below
#Used QuickGO to search and fill in the remaining NA parent terms https://www.ebi.ac.uk/QuickGO/term/GO:1901363
#DMG go plot
```{r}
#upload complete parent term dataframe DMGs and make figures

#upload complete go enrichment summary for DMGs
DMG.complete <- read.csv("RAnalysis/Data/WGBS/DMG_GO_parent_terms_complete.10x.csv", header = TRUE)

#adjust p-values 
DMG.complete$bh_adjust <-  p.adjust(DMG.complete$over_represented_pvalue, method="BH") 

GO.plot.DMG.complete <-  ggplot(DMG.complete, aes(x = ontology, y = term)) + 
    geom_point(aes(size=numInCat, fill=bh_adjust, colour=bh_adjust)) + 
    scale_size(name="Number of Genes", range=c(5,12), limits=c(1,2120))+
    scale_fill_gradient(name = "Adjusted p-value", low="black", high="gray")+
    scale_colour_gradient(name = "Adjusted p-value", low="black", high="gray")+
    xlab("Ontology Category")+
    facet_grid(slim_term ~ ., scales = "free", space='free', labeller = label_wrap_gen(width = 55, multi_line = TRUE))+
    ggtitle("")+
    theme_bw() + 
    theme(panel.border = element_blank(), 
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"),
          strip.background =element_rect(fill="white"),
          strip.text.y = element_text(angle=0, size = 18, face="bold"),
          strip.text.x = element_text(size = 16),
          axis.text.x = element_text(size = 20, angle=45, hjust=1),
          axis.text.y = element_text(size = 16),
          axis.title.x = element_text(size = 22, face="bold"),
          legend.text=element_text(size=16),
          title=element_text(size=22, face="bold"),
          plot.title = element_text(hjust = 0),
          axis.title.y = element_blank())

 
ggsave("RAnalysis/Output/WGBS/DMG.complete.slimterms.pdf", GO.plot.DMG.complete, width = 20, height = 49.5, units = c("in"))
ggsave("RAnalysis/Output/WGBS/DMG.complete.slimterms.png", GO.plot.DMG.complete, width = 20, height = 49.5, units = c("in"))


```


```{r}
#upload complete parent term dataframe DMGs and make figures

#upload complete go enrichment summary for DMGs
DEG.complete <- read.csv("RAnalysis/Data/RNA-seq/P.verrucosa/Host/DEG_GO_parent_terms_complete.csv", header = TRUE)

#add column to specific type of measure
#DEG.complete <- DEG.complete %>% add_column(type = "DEG.GO.terms")

#adjust p-values 
DEG.complete$bh_adjust <-  p.adjust(DEG.complete$over_represented_pvalue, method="BH") 

GO.plot.DEG.complete <-  ggplot(DEG.complete, aes(x = ontology, y = term)) + 
    geom_point(aes(size=numInCat, fill=bh_adjust, colour=bh_adjust)) + 
    scale_size(name="Number of Genes", range=c(5,12), limits=c(1,550))+
    scale_fill_gradient(name = "Adjusted p-value", low="black", high="gray")+
    scale_colour_gradient(name = "Adjusted p-value", low="black", high="gray")+
    xlab("Ontology Category")+
    facet_grid(slim_term ~ ., scales = "free", space='free', labeller = label_wrap_gen(width = 55, multi_line = TRUE))+
    ggtitle("")+
    theme_bw() + 
    theme(panel.border = element_blank(), 
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"),
          strip.background =element_rect(fill="white"),
          strip.text.y = element_text(angle=0, size = 18, face="bold"),
          strip.text.x = element_text(size = 16),
          axis.text.x = element_text(size = 20, angle=45, hjust=1),
          axis.text.y = element_text(size = 16),
          axis.title.x = element_text(size = 22, face="bold"),
          legend.text=element_text(size=16),
          title=element_text(size=22, face="bold"),
          plot.title = element_text(hjust = 0),
          axis.title.y = element_blank())

 
ggsave("RAnalysis/Output/Final_Figures/Host/DEG.complete.slimterms.pdf", GO.plot.DEG.complete , width = 25, height = 49, units = c("in"))


```


#load png image of DEG PCA and DMG PCA to combine for each figure
```{r}
DEG.PCA <- readPNG("RAnalysis/Output/Final_Figures/Host/DEGSeq2_DEG_PCA_plot_host.png")
PCA.DEG <- rasterGrob(DEG.PCA, interpolate=TRUE)

DMG.PCA <- readPNG("RAnalysis/Output/Final_Figures/Host/DMG_PCA_plot_host.png")
PCA.DMG <- rasterGrob(DMG.PCA, interpolate=TRUE)

```

#make DEG plot figure 2 manuscript
```{r}

#filter for numincat sum per term and remove CC ontology for dataframe before making plot BP
filter.DEG.BP <-DEG.complete %>%
  filter(ontology %in% c("BP"))

filt.sum.DEG.BP <- filter.DEG.BP %>% group_by(slim_term) %>% 
  summarise(sum_num=sum(numInCat))

filt.mean.DEG.BP <- filter.DEG.BP %>% group_by(slim_term) %>% 
  summarise(pvalue=mean(bh_adjust))

subset.DEG.BP <- left_join(filt.mean.DEG.BP, filt.sum.DEG.BP)

subset.DEG.BP <- subset.DEG.BP %>%
  mutate(ontology = "Biological Process")

GO.plot.DEG.BP <-  ggplot(subset.DEG.BP, aes(x = ontology, y = slim_term)) + 
    geom_point(aes(size=sum_num, fill=pvalue, colour=pvalue)) + 
    scale_size(name="Number of Genes", range=c(5,12), limits=c(1,800))+
    scale_fill_gradient(name = "Adjusted p-value", low="black", high="gray")+
    scale_colour_gradient(name = "Adjusted p-value", low="black", high="gray")+
    xlab("")+
    facet_grid(slim_term ~ ., scales = "free", space='free', labeller = label_wrap_gen(width = 55, multi_line = TRUE), switch="y")+
    ggtitle("")+
    theme_bw() + 
    #scale_y_discrete(limits=rev) +
    theme(strip.text.y.left = element_text(angle = 0),
          panel.border = element_blank(), 
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"),
          strip.background =element_rect(fill="white"),
          strip.text.y = element_text(angle=0, size = 18, face="bold"),
          strip.text.x = element_text(size = 30, face="bold"),
          axis.text.x = element_text(size = 30, face="bold"),
          axis.text.y=element_blank(),
          axis.ticks.y = element_blank(),
          axis.title.x = element_text(size = 30, face="bold"),
          legend.text=element_text(size=30, face="bold"),
          title=element_text(size=30, face="bold"),
          plot.title = element_text(hjust = 0, face="bold"),
          axis.title.y = element_blank())

ggsave("RAnalysis/Output/RNA-seq/DEG/Host/DEG.BP.png", GO.plot.DEG.BP, width = 20, height = 40, units = c("in"))
ggsave("RAnalysis/Output/RNA-seq/DEG/Host/DEG.BP.pdf", GO.plot.DEG.BP, width = 20, height = 20, units = c("in"))


#filter for numincat sum per term and remove CC ontology for dataframe before making plot BP
filter.DEG.MF <-DEG.complete %>%
  filter(ontology %in% c("MF"))

filt.sum.DEG.MF <- filter.DEG.MF%>% group_by(slim_term) %>% 
  summarise(sum_num=sum(numInCat))

filt.mean.DEG.MF <- filter.DEG.MF %>% group_by(slim_term) %>% 
  summarise(pvalue=mean(bh_adjust))

subset.DEG.MF <- left_join(filt.mean.DEG.MF, filt.sum.DEG.MF)

subset.DEG.MF <- subset.DEG.MF %>%
  mutate(ontology = "Molecular Function")

GO.plot.DEG.MF <-  ggplot(subset.DEG.MF, aes(x = ontology, y = slim_term)) + 
    geom_point(aes(size=sum_num, fill=pvalue, colour=pvalue)) + 
    scale_size(name="Number of Genes", range=c(5,12), limits=c(1,1455))+
    scale_fill_gradient(name = "Adjusted p-value", low="black", high="gray")+
    scale_colour_gradient(name = "Adjusted p-value", low="black", high="gray")+
    xlab("")+
    facet_grid(slim_term ~ ., scales = "free", space='free', labeller = label_wrap_gen(width = 55, multi_line = TRUE), switch="y")+
    ggtitle("")+
    theme_bw() + 
    #scale_y_discrete(limits=rev) +
    theme(strip.text.y.left = element_text(angle = 0),
          panel.border = element_blank(), 
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"),
          strip.background =element_rect(fill="white"),
          strip.text.y = element_text(angle=0, size = 18, face="bold"),
          strip.text.x = element_text(size = 30, face="bold"),
          axis.text.x = element_text(size = 30, face="bold"),
          axis.text.y=element_blank(),
          axis.ticks.y = element_blank(),
          axis.title.x = element_text(size = 30, face="bold"),
          legend.text=element_text(size=30, face="bold"),
          title=element_text(size=30, face="bold"),
          plot.title = element_text(hjust = 0, face="bold"),
          axis.title.y = element_blank())

ggsave("RAnalysis/Output/RNA-seq/DEG/Host/DEG.MF.png", GO.plot.DEG.MF, width = 10, height = 10, units = c("in"))
ggsave("RAnalysis/Output/RNA-seq/DEG/Host/DEG.MF.pdf", GO.plot.DEG.MF, width = 10, height = 10, units = c("in"))

```

#make DMG figure 3 for manuscript 
```{r}

#filter for numincat sum per term and remove CC ontology for dataframe before making plot BP
filter.DMG.BP <-DMG.complete %>%
  filter(ontology %in% c("BP"))

filt.sum.BP <- filter.DMG.BP %>% group_by(slim_term) %>% 
  summarise(sum_num=sum(numInCat))

filt.mean.BP <- filter.DMG.BP %>% group_by(slim_term) %>% 
  summarise(pvalue=mean(bh_adjust))

subset.BP <- left_join(filt.mean.BP, filt.sum.BP)

subset.BP <- subset.BP %>%
  mutate(ontology = "Biological Process")

GO.plot.DMG.BP <-  ggplot(subset.BP, aes(x = ontology, y = slim_term)) + 
    geom_point(aes(size=sum_num, fill=pvalue, colour=pvalue)) + 
    scale_size(name="Number of Genes", range=c(5,12), limits=c(1,1528))+
    scale_fill_gradient(name = "Adjusted p-value", low="black", high="gray")+
    scale_colour_gradient(name = "Adjusted p-value", low="black", high="gray")+
    xlab("")+
    facet_grid(slim_term ~ ., scales = "free", space='free', labeller = label_wrap_gen(width = 55, multi_line = TRUE), switch="y")+
    ggtitle("")+
    theme_bw() + 
    #scale_y_discrete(limits=rev) +
    theme(strip.text.y.left = element_text(angle = 0),
          panel.border = element_blank(), 
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"),
          strip.background =element_rect(fill="white"),
          strip.text.y = element_text(angle=0, size = 18, face="bold"),
          strip.text.x = element_text(size = 30, face="bold"),
          axis.text.x = element_text(size = 30, face="bold"),
          axis.text.y=element_blank(),
          axis.ticks.y = element_blank(),
          axis.title.x = element_text(size = 30, face="bold"),
          legend.text=element_text(size=30, face="bold"),
          title=element_text(size=30, face="bold"),
          plot.title = element_text(hjust = 0, face="bold"),
          axis.title.y = element_blank())

ggsave("RAnalysis/Output/WGBS/DMG.BP.png", GO.plot.DMG.BP, width = 20, height = 35, units = c("in"))
ggsave("RAnalysis/Output/WGBS/DMG.BP.pdf", GO.plot.DMG.BP, width = 20, height = 20, units = c("in"))


#filter for numincat sum per term and remove CC ontology for dataframe before making plot BP
filter.DMG.MF <-DMG.complete %>%
  filter(ontology %in% c("MF"))

filt.sum.MF <- filter.DMG.MF%>% group_by(slim_term) %>% 
  summarise(sum_num=sum(numInCat))

filt.mean.MF <- filter.DMG.MF %>% group_by(slim_term) %>% 
  summarise(pvalue=mean(bh_adjust))

subset.MF <- left_join(filt.mean.MF, filt.sum.MF)

subset.MF <- subset.MF %>%
  mutate(ontology = "Molecular Function")

GO.plot.DMG.MF <-  ggplot(subset.MF, aes(x = ontology, y = slim_term)) + 
    geom_point(aes(fill=pvalue, colour=pvalue, size=sum_num)) + 
    scale_fill_gradient(name = "Adjusted p-value", low="black", high="gray")+
    scale_colour_gradient(name = "Adjusted p-value", low="black", high="gray")+
    scale_size(name="Number of Genes", range=c(5,12), limits=c(1,5600))+
    xlab("")+
    facet_grid(slim_term ~ ., scales = "free", space='free', labeller = label_wrap_gen(width = 55, multi_line = TRUE), switch="y")+
    ggtitle("")+
    theme_bw() + 
    #scale_y_discrete(limits=rev) +
    theme(strip.text.y.left = element_text(angle = 0),
          panel.border = element_blank(), 
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"),
          strip.background =element_rect(fill="white"),
          strip.text.y = element_text(angle=0, size = 18, face="bold"),
          strip.text.x = element_text(size = 30, face="bold"),
          axis.text.x = element_text(size = 30, face="bold"),
          axis.text.y=element_blank(),
          axis.ticks.y = element_blank(),
          axis.title.x = element_text(size = 30, face="bold"),
          legend.text=element_text(size=30, face="bold"),
          title=element_text(size=30, face="bold"),
          plot.title = element_text(hjust = 0, face="bold"),
          axis.title.y = element_blank())

ggsave("RAnalysis/Output/WGBS/DMG.MF.png", GO.plot.DMG.MF, width = 10, height = 10, units = c("in"))
ggsave("RAnalysis/Output/WGBS/DMG.MF.pdf", GO.plot.DMG.MF, width = 10, height = 10, units = c("in"))



```

#use patchwork to make figure 2 and 3 wiht PCA as a, BP as B and MF as C for DEG and DMG

```{r}

bottom_row.DEG <- plot_grid(GO.plot.DEG.BP, GO.plot.DEG.MF, label_size = 60)
Figure2 <- plot_grid(PCA.DEG, bottom_row.DEG, labels = c('A', ''), label_size = 60, ncol = 1) 

ggsave(filename=paste0("RAnalysis/Output/Final_Figures/Host/manuscript/DEG.png"), plot=Figure2, dpi=300,width=50, height=60, units="in", limitsize = FALSE)
ggsave(filename=paste0("RAnalysis/Output/Final_Figures/Host/manuscript/DEG.pdf"), plot=Figure2, dpi=300, width=50, height=60, units="in", limitsize = FALSE)

bottom_row.DMG <- plot_grid(GO.plot.DMG.BP, GO.plot.DMG.MF, label_size = 60)
Figure3 <- plot_grid(PCA.DMG, bottom_row.DMG, labels = c('A', ''), label_size = 60, ncol = 1)

ggsave(filename=paste0("RAnalysis/Output/Final_Figures/Host/manuscript/DMG.png"), plot=Figure3, dpi=300,width=50, height=60, units="in", limitsize = FALSE)
ggsave(filename=paste0("RAnalysis/Output/Final_Figures/Host/manuscript/DMG.pdf"), plot=Figure3, dpi=300, width=50, height=60, units="in", limitsize = FALSE)
```



#facet by ontology, x axis of dot plots could be genes and methylation to see which are shared and each ontology level what is overlap in each
#see next is how gene and meth functions overlap facet grouping by parent term and which gives you overlap in parent terms
#DEG go plot
```{r}
#add column for deg dmg specific category to each dataframe
DMG.complete <- DMG.complete %>% add_column(type = "DMG.GO.terms")

#rename column from genes to gene
DEG.complete <- DEG.complete %>% 
       dplyr::rename("GOterm" = "category")

#rename column from genes to gene
DMG.complete <- DMG.complete %>% 
       dplyr::rename("GOterm" = "category")


#combine two dataframes to have both DEG and DMG terms in same data frame
DEG.DMG <- rbind(DMG.complete, DEG.complete)

#filter for just the biological process ontology
DEG.DMG.BP <- DEG.DMG %>%
filter(ontology == "BP")

#output csv of shared DMG and DEG terms
write.csv(DEG.DMG, "RAnalysis/Output/meth_express_compare/global.DEG.DMG.goterms.csv")

BP.df <- na.omit(DEG.DMG.BP)


#make biological process DEG DMG comparison figure
GO.plot.BP.DEG.DMG <-  ggplot(BP.df, aes(x = type, y = term)) + 
    geom_point(aes(size=numInCat, fill=bh_adjust, colour=bh_adjust)) + 
    scale_size(name="Number of Genes", range=c(5,12), limits=c(1,550))+
    scale_fill_gradient(name = "Adjusted p-value", low="black", high="gray")+
    scale_colour_gradient(name = "Adjusted p-value", low="black", high="gray")+
    xlab("Method")+
    facet_grid(slim_term ~ ., scales = "free", space='free', labeller = label_wrap_gen(width = 80, multi_line = TRUE))+
    ggtitle("Biological Process")+
    theme_bw() + 
    theme(panel.border = element_blank(), 
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"),
          strip.background =element_rect(fill="white"),
          strip.text.y = element_text(angle=0, size = 18, face="bold"),
          strip.text.x = element_text(size = 16),
          axis.text.x = element_text(size = 20, angle=45, hjust=1),
          axis.text.y = element_text(size = 16),
          axis.title.x = element_text(size = 22, face="bold"),
          legend.text=element_text(size=16),
          title=element_text(size=22, face="bold"),
          plot.title = element_text(hjust = 0.5),
          axis.title.y = element_blank());GO.plot.BP.DEG.DMG

 

ggsave("RAnalysis/Output/Final_Figures/Host/GO.BP.DEG.DMG.pdf", GO.plot.BP.DEG.DMG , width = 35, height = 49, units = c("in"))

```


#make final DMG DEG molecular function comparison plot
```{r}
#filter for just the molecular function ontology
DEG.DMG.MF <- DEG.DMG %>%
filter(ontology == "MF")

#make biological process DEG DMG comparison figure
GO.plot.MF.DEG.DMG <-  ggplot(DEG.DMG.MF, aes(x = type, y = term)) + 
    geom_point(aes(size=numInCat, fill=bh_adjust, colour=bh_adjust)) + 
    scale_size(name="Number of Genes", range=c(5,12), limits=c(1,550))+
    scale_fill_gradient(name = "Adjusted p-value", low="black", high="gray")+
    scale_colour_gradient(name = "Adjusted p-value", low="black", high="gray")+
    xlab("Method")+
    facet_grid(slim_term ~ ., scales = "free", space='free', labeller = label_wrap_gen(width = 55, multi_line = TRUE))+
    ggtitle("Molecular Function")+
    theme_bw() + 
    theme(panel.border = element_blank(), 
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"),
          strip.background =element_rect(fill="white"),
          strip.text.y = element_text(angle=0, size = 18, face="bold"),
          strip.text.x = element_text(size = 16),
          axis.text.x = element_text(size = 20, angle=45, hjust=1),
          axis.text.y = element_text(size = 16),
          axis.title.x = element_text(size = 22, face="bold"),
          legend.text=element_text(size=16),
          title=element_text(size=22, face="bold"),
          plot.title = element_text(hjust = 0.5),
          axis.title.y = element_blank());GO.plot.MF.DEG.DMG

 

ggsave("RAnalysis/Output/Final_Figures/Host/GO.MF.DEG.DMG.pdf", GO.plot.MF.DEG.DMG , width = 20, height = 49, units = c("in"))

```


#make final DMG DEG cellular component comparison plot
```{r}
#filter for just the molecular function ontology
DEG.DMG.CC <- DEG.DMG %>%
filter(ontology == "CC")

#make biological process DEG DMG comparison figure
GO.plot.CC.DEG.DMG <-  ggplot(DEG.DMG.CC, aes(x = type, y = term)) + 
    geom_point(aes(size=numInCat, fill=bh_adjust, colour=bh_adjust)) + 
    scale_size(name="Number of Genes", range=c(5,12), limits=c(1,550))+
    scale_fill_gradient(name = "Adjusted p-value", low="black", high="gray")+
    scale_colour_gradient(name = "Adjusted p-value", low="black", high="gray")+
    xlab("Method")+
    facet_grid(slim_term ~ ., scales = "free", space='free', labeller = label_wrap_gen(width = 55, multi_line = TRUE))+
    ggtitle("Cellular Component")+
    theme_bw() + 
    theme(panel.border = element_blank(), 
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"),
          strip.background =element_rect(fill="white"),
          strip.text.y = element_text(angle=0, size = 18, face="bold"),
          strip.text.x = element_text(size = 16),
          axis.text.x = element_text(size = 20, angle=45, hjust=1),
          axis.text.y = element_text(size = 16),
          axis.title.x = element_text(size = 22, face="bold"),
          legend.text=element_text(size=16),
          title=element_text(size=22, face="bold"),
          plot.title = element_text(hjust = 0.5),
          axis.title.y = element_blank());GO.plot.CC.DEG.DMG

 

ggsave("RAnalysis/Output/Final_Figures/Host/GO.CC.DEG.DMG.pdf", GO.plot.CC.DEG.DMG , width = 20, height = 30, units = c("in"))

```

#compare gene lists for correlation, DEG, DMG need to bring in statistic data frame
```{r}
#methylation and expression dataframe calculated and complete in meth_GE_compare.Rmd script, found here: https://github.com/hputnam/Becker_E5/blob/master/RAnalysis/Scripts/methyl_expression_compare/meth_GE_compare.Rmd
#this data frame lists only the genes found in meth and express data frames
meth.express <- readRDS("RAnalysis/Data/GE_meth_compare/Multi_geneSummaryComplete.RData")

#remove _gene_ in gene id to match other data frames
meth.express$gene_id <-  gsub("_gene_","_", meth.express$gene_id)

#this data frame lists all methylated genes found 
methyl.annot <- readRDS("RAnalysis/Data/GE_meth_compare/meth_annot_full.RData")

#remove _gene_ in gene id to match other data frames
methyl.annot$gene_id <-  gsub("_gene_","_", methyl.annot$gene_id)

#this data frame lists all expressed genes found
express.annot <- readRDS("RAnalysis/Data/GE_meth_compare/gene_annot_full.RData")

#remove _gene_ in gene id to match other data frames
express.annot$gene_id <-  gsub("_gene_","_", express.annot$gene_id)

#remove gene_ from column 2 to 8
colnames(express.annot)[2:8] <-  sub(".*gene_", "", colnames(express.annot)[2:8])

```


#Filter dataframes for gene ontology groups that are involved in nitrogen pathways, stress pathways to see if expression and methylation are correlated
```{r}
#filter by any category that contained the word, could do anything that has the word nitrogen, stress, and heat 

vector1 <- c("nitrogen compound metabolic process", "response to nitrogen compound", "response to nutrient levels", "heat shock protein binding", "response to stress", "cellular nitrogen compound", "glutamate", "glutamine", "protein metabolic proces", "amino acid")

phenotype.pathways.DEG <-DEG.complete%>%
      filter(grepl(paste(vector1, collapse="|"), paste(slim_term, term))) 

phenotype.pathways.DMG <- DMG.complete %>% 
      filter(grepl(paste(vector1, collapse="|"), paste(slim_term, term))) 

```

#change data frames from wide to long format
```{r}
#change data frames from wide to long so each gene has a row 
long.pathway.DEG <- phenotype.pathways.DEG %>% 
    mutate(genes = strsplit(as.character(genes), ",")) %>% #defines what character separates the list
    unnest(genes) #unnests and makes each gene have its own row

#rename column from genes to gene
long.pathway.DEG <- long.pathway.DEG %>% 
       dplyr::rename("gene_id" = "genes")

#remove type column from data frame to match DMG 
long.pathway.DEG = subset(long.pathway.DEG, select = -c(type) )

#remove extra first space in gene_id names in column 
long.pathway.DEG$gene_id <- gsub(' Pver', 'Pver', long.pathway.DEG$gene_id)

#change data frames from wide to long so each gene has a row
long.pathway.DMG <- phenotype.pathways.DMG %>% 
    mutate(genes = strsplit(as.character(genes), ",")) %>% 
    unnest(genes)

#remove the _gene_ in the DMG gene list to match DEG gene numbers
long.pathway.DMG$genes <-  gsub("_gene_","_", long.pathway.DMG$genes)

#rename column from genes to gene
long.pathway.DMG <- long.pathway.DMG %>% 
       dplyr::rename("gene_id" = "genes")

#remove extra first space in gene_id names in column 
long.pathway.DMG$gene_id <- gsub(' Pver', 'Pver', long.pathway.DMG$gene_id)

```

# Combine DMG and DEG data frames, by binding below so all rows are present
```{r}
#add column to specific type of measure
long.pathway.DEG <- long.pathway.DEG %>% add_column(type = "DEG.GO.terms")

#DMG columns = 12 rows and DEG = 435 rows, total rows should be 447
DMG.DEG.dat <- rbind(long.pathway.DEG, long.pathway.DMG)
#yes rows equal estimate

#write hypothesis driven csv
write.csv(DMG.DEG.dat, "RAnalysis/Output/meth_express_compare/hypothesis.DEG.DMG.goterms.csv")
```


```{r}
#combine all meth express data frame with DMG.DEG data frame
#this will merge only genes found in both DMG and DEG
meth.exp.filt <- merge(meth.express, DMG.DEG.dat, by = 'gene_id')

#this will merge to list all DEG and DMG express and meth statistics
meth.merge <- merge(methyl.annot, long.pathway.DMG)

express.merge <- merge(express.annot, long.pathway.DEG, by = 'gene_id')

#the reason some expressed or methylated genes do not merge with stats are due to filtering for >3 CpG in methylation and >0 mean expression


```


# Look at correlation for mean methylation and gene expression in hypothesis driven approach
```{r}
## Mean Methylation vs Gene Expression following Liew et al
df <- meth.exp.filt

df %>% drop_na()

#control log gene expression x % methylation
df$log_gene <- log(df$gene_mean_express+1)
df$per_meth <- df$mean_meth * 100

#make plot with correlation tests
#decide between pearson, kendall, or spearman
#first test normality
shapiro.test(df$gene_mean_express) # <0.05
shapiro.test(df$mean_meth) # <0.05
#both variables are less then 0.05 indicating data do not follow normal distribution

#use kendall test for correlation, Kendall’s is more robust and generally the preferred method between spearman and kendall

meth.express.specific <- ggplot(df,aes(x=per_meth, y=log_gene,)) + 
    geom_point() + 
    geom_smooth(method=lm,colour="orange",size=2,na.rm = TRUE) +
    scale_color_viridis() +
    labs(x="DNA Methylation (%)", y = "log10 (mean expression +1)") +
  stat_cor(method = "kendall", digits = 3, label.x = 60, label.y = 0.5 ) +
    theme_cowplot();meth.express.specific

### Figure Mean Methylation vs Gene Expression###

ggsave(meth.express.specific, filename="RAnalysis/Output/meth_express_compare/pathway.specific.express.meth.png", height = 7, width = 10)
```


```{r}
#find unique GO slim terms that overlap between DEG and DMG lists
#DEG
unique.DEG <- DEG.complete %>%
  distinct(slim_term, ontology) %>%
  tibble::as_tibble()  # Convert to tibble

#filter NAs
unique.DEG <- unique.DEG %>%
  filter(!is.na(slim_term) & !is.na(ontology))

#DMG
unique.DMG <- DMG.complete %>%
  distinct(slim_term, ontology) %>%
  tibble::as_tibble()  # Convert to tibble

#filter NAs
unique.DMG <- unique.DMG %>%
  filter(!is.na(slim_term) & !is.na(ontology))

#merge to show overlapping terms
overlap <- merge(unique.DEG, unique.DMG, on="slim_term")
#32 unique terms

#make biological process DEG DMG comparison figure
hypo.DEG.DMG <-  ggplot(df, aes(x = type, y = term)) + 
    geom_point(aes(size=numInCat, fill=bh_adjust, colour=bh_adjust)) + 
    scale_size(name="Number of Genes", range=c(5,12), limits=c(1,2000))+
    scale_fill_gradient(name = "Adjusted p-value", low="black", high="gray")+
    scale_colour_gradient(name = "Adjusted p-value", low="black", high="gray")+
    xlab("Method")+
    facet_grid(slim_term ~ ., scales = "free", space='free', labeller = label_wrap_gen(width = 55, multi_line = TRUE))+
    #ggtitle()+
    theme_bw() + 
    theme(panel.border = element_blank(), 
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"),
          strip.background =element_rect(fill="white"),
          strip.text.y = element_text(angle=0, size = 18, face="bold"),
          strip.text.x = element_text(size = 16),
          axis.text.x = element_text(size = 20, angle=45, hjust=1),
          axis.text.y = element_text(size = 16),
          axis.title.x = element_text(size = 22, face="bold"),
          legend.text=element_text(size=16),
          title=element_text(size=22, face="bold"),
          plot.title = element_text(hjust = 0.5),
          axis.title.y = element_blank());hypo.DEG.DMG

ggsave("RAnalysis/Output/Final_Figures/Host/GO.hypo.DEG.DMG.pdf", hypo.DEG.DMG , width = 20, height = 30, units = c("in"))
ggsave("RAnalysis/Output/Final_Figures/Host/GO.hypo.DEG.DMG.png", hypo.DEG.DMG , width = 20, height = 30, units = c("in"))


```

#make dataframes and figures for specific hypothesis driven pathways and responses from DEGs and DMGs

#nutrient response
```{r}

nutrient <- c("nutrient")

nutrient.DEG <-DEG.complete%>%
      filter(grepl(paste(nutrient, collapse="|"), paste(slim_term, term))) 

nutrient.DMG <- DMG.complete %>% 
      filter(grepl(paste(nutrient, collapse="|"), paste(slim_term, term)))

nutrient.dat <- rbind(nutrient.DEG,nutrient.DMG)

  #plot significantly enriched GO terms by Slim Category faceted by slim term 
  nutrient.GO.plot <-  ggplot(nutrient.dat, aes(x = type, y = term)) + 
    geom_point(aes(size=numInCat, fill=bh_adjust, colour=bh_adjust)) + 
    scale_size(name="Number of Genes", range=c(5,12), limits=c(1,50))+
    scale_fill_gradient(name = "Adjusted p-value", low="black", high="gray")+
    scale_colour_gradient(name = "Adjusted p-value", low="black", high="gray")+
    facet_grid(slim_term ~ ., scales = "free", space='free', labeller = label_wrap_gen(width = 6, multi_line = TRUE))+
    xlab("Analysis")+
    ggtitle("Nutrient Response")+
    theme_bw() + 
    theme(panel.border = element_blank(), 
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"),
          strip.background =element_rect(fill="white"),
          strip.text.y = element_text(angle=0, size = 18, face="bold"),
          strip.text.x = element_text(size = 16),
          axis.text.x = element_text(size = 20, angle=45, hjust=1),
          axis.text.y = element_text(size = 16),
          axis.title.x = element_text(size = 22, face="bold"),
          legend.text=element_text(size=16),
          title=element_text(size=22, face="bold"),
          plot.title = element_text(hjust = 0.6),
          axis.title.y = element_blank())
  
ggsave(filename=paste0("RAnalysis/Output/Final_Figures/Host/specific_pathways/nutrient.DMG.DEG.png"), plot=nutrient.GO.plot, dpi=300, width=15, height=10, units="in")
ggsave(filename=paste0("RAnalysis/Output/Final_Figures/Host/specific_pathways/nutrient.DMG.DEG.pdf"), plot=nutrient.GO.plot, dpi=300, width=15, height=10, units="in")

```

#nitrogen metabolism
```{r}

nitrogen <- c("nitrogen")

nitrogen.DEG <-DEG.complete%>%
      filter(grepl(paste(nitrogen, collapse="|"), paste(slim_term, term))) 

nitrogen.DMG <- DMG.complete %>% 
      filter(grepl(paste(nitrogen, collapse="|"), paste(slim_term, term)))

nitrogen.dat <- rbind(nitrogen.DEG,nitrogen.DMG)

  #plot significantly enriched GO terms by Slim Category faceted by slim term 
  nitrogen.GO.plot <-  ggplot(nitrogen.dat, aes(x = type, y = term)) + 
    geom_point(aes(size=numInCat, fill=bh_adjust, colour=bh_adjust)) + 
    scale_size(name="Number of Genes", range=c(5,12), limits=c(1,100))+
    scale_fill_gradient(name = "Adjusted p-value", low="black", high="gray")+
    scale_colour_gradient(name = "Adjusted p-value", low="black", high="gray")+
    facet_grid(slim_term ~ ., scales = "free", space='free', labeller = label_wrap_gen(width = 8, multi_line = TRUE))+
    xlab("Analysis")+
    ggtitle("Nitrogen Metabolism")+
    theme_bw() + 
    theme(panel.border = element_blank(), 
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"),
          strip.background =element_rect(fill="white"),
          strip.text.y = element_text(angle=0, size = 18, face="bold"),
          strip.text.x = element_text(size = 16),
          axis.text.x = element_text(size = 20, angle=45, hjust=1),
          axis.text.y = element_text(size = 16),
          axis.title.x = element_text(size = 22, face="bold"),
          legend.text=element_text(size=16),
          title=element_text(size=22, face="bold"),
          plot.title = element_text(hjust = 0.6),
          axis.title.y = element_blank())
  
ggsave(filename=paste0("RAnalysis/Output/Final_Figures/Host/specific_pathways/nitrogen.DMG.DEG.png"), plot=nitrogen.GO.plot, dpi=300, width=15, height=15, units="in")
ggsave(filename=paste0("RAnalysis/Output/Final_Figures/Host/specific_pathways/nitrogen.DMG.DEG.pdf"), plot=nitrogen.GO.plot, dpi=300, width=15, height=15, units="in")

```

#heat stress response
```{r}

heat <- c("heat", "DNA replication", "DNA repair", "immune", "protein folding", "stress", "DNA")

heat.DEG <-DEG.complete%>%
      filter(grepl(paste(heat, collapse="|"), paste(slim_term, term))) 

heat.DMG <- DMG.complete %>% 
      filter(grepl(paste(heat, collapse="|"), paste(slim_term, term)))

heat.dat <- rbind(heat.DEG,heat.DMG)

  #plot significantly enriched GO terms by Slim Category faceted by slim term 
  heat.GO.plot <-  ggplot(heat.dat, aes(x = type, y = term)) + 
    geom_point(aes(size=numInCat, fill=bh_adjust, colour=bh_adjust)) + 
    scale_size(name="Number of Genes", range=c(5,12), limits=c(1,1500))+
    scale_fill_gradient(name = "Adjusted p-value", low="black", high="gray")+
    scale_colour_gradient(name = "Adjusted p-value", low="black", high="gray")+
    facet_grid(slim_term ~ ., scales = "free", space='free', labeller = label_wrap_gen(width = 50, multi_line = TRUE))+
    xlab("Analysis")+
    ggtitle("Heat Stress Response")+
    theme_bw() + 
    theme(panel.border = element_blank(), 
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"),
          strip.background =element_rect(fill="white"),
          strip.text.y = element_text(angle=0, size = 18, face="bold"),
          strip.text.x = element_text(size = 16),
          axis.text.x = element_text(size = 20, angle=45, hjust=1),
          axis.text.y = element_text(size = 16),
          axis.title.x = element_text(size = 22, face="bold"),
          legend.text=element_text(size=16),
          title=element_text(size=22, face="bold"),
          plot.title = element_text(hjust = 0.6),
          axis.title.y = element_blank())
  
ggsave(filename=paste0("RAnalysis/Output/Final_Figures/Host/specific_pathways/heat.DMG.DEG.png"), plot=heat.GO.plot, dpi=300, width=22, height=15, units="in")
ggsave(filename=paste0("RAnalysis/Output/Final_Figures/Host/specific_pathways/heat.DMG.DEG.pdf"), plot=heat.GO.plot, dpi=300, width=22, height=15, units="in")


```


#carbohydrate metabolism
```{r}

carb <- c("carbohydrate")

carb.DEG <-DEG.complete%>%
      filter(grepl(paste(carb, collapse="|"), paste(slim_term, term))) 

carb.DMG <- DMG.complete %>% 
      filter(grepl(paste(carb, collapse="|"), paste(slim_term, term)))

carb.dat <- rbind(carb.DEG,carb.DMG)

  #plot significantly enriched GO terms by Slim Category faceted by slim term 
  carb.GO.plot <-  ggplot(carb.dat, aes(x = type, y = term)) + 
    geom_point(aes(size=numInCat, fill=bh_adjust, colour=bh_adjust)) + 
    scale_size(name="Number of Genes", range=c(5,12), limits=c(1,50))+
    scale_fill_gradient(name = "Adjusted p-value", low="black", high="gray")+
    scale_colour_gradient(name = "Adjusted p-value", low="black", high="gray")+
    facet_grid(slim_term ~ ., scales = "free", space='free', labeller = label_wrap_gen(width = 8, multi_line = TRUE))+
    xlab("Analysis")+
    ggtitle("Carbohydrate Metabolism")+
    theme_bw() + 
    theme(panel.border = element_blank(), 
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"),
          strip.background =element_rect(fill="white"),
          strip.text.y = element_text(angle=0, size = 18, face="bold"),
          strip.text.x = element_text(size = 16),
          axis.text.x = element_text(size = 20, angle=45, hjust=1),
          axis.text.y = element_text(size = 16),
          axis.title.x = element_text(size = 22, face="bold"),
          legend.text=element_text(size=16),
          title=element_text(size=22, face="bold"),
          plot.title = element_text(hjust = 0.6),
          axis.title.y = element_blank())
  
ggsave(filename=paste0("RAnalysis/Output/Final_Figures/Host/specific_pathways/carb.DMG.DEG.png"), plot=carb.GO.plot, dpi=300, width=15, height=15, units="in")
ggsave(filename=paste0("RAnalysis/Output/Final_Figures/Host/specific_pathways/carb.DMG.DEG.pdf"), plot=carb.GO.plot, dpi=300, width=15, height=15, units="in")
 

```

#amino acid metabolism
```{r}

amino <- c("amino acid", "amino")

amino.DEG <-DEG.complete%>%
      filter(grepl(paste(amino, collapse="|"), paste(slim_term, term))) 

amino.DMG <- DMG.complete %>% 
      filter(grepl(paste(amino, collapse="|"), paste(slim_term, term)))

amino.dat <- rbind(amino.DEG,amino.DMG)

  #plot significantly enriched GO terms by Slim Category faceted by slim term 
  amino.GO.plot <-  ggplot(amino.dat, aes(x = type, y = term)) + 
    geom_point(aes(size=numInCat, fill=bh_adjust, colour=bh_adjust)) + 
    scale_size(name="Number of Genes", range=c(5,12), limits=c(1,35))+
    scale_fill_gradient(name = "Adjusted p-value", low="black", high="gray")+
    scale_colour_gradient(name = "Adjusted p-value", low="black", high="gray")+
    facet_grid(slim_term ~ ., scales = "free", space='free', labeller = label_wrap_gen(width = 8, multi_line = TRUE))+
    xlab("Analysis")+
    ggtitle("Amino Acid Metabolism")+
    theme_bw() + 
    theme(panel.border = element_blank(), 
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"),
          strip.background =element_rect(fill="white"),
          strip.text.y = element_text(angle=0, size = 18, face="bold"),
          strip.text.x = element_text(size = 16),
          axis.text.x = element_text(size = 20, angle=45, hjust=1),
          axis.text.y = element_text(size = 16),
          axis.title.x = element_text(size = 22, face="bold"),
          legend.text=element_text(size=16),
          title=element_text(size=22, face="bold"),
          plot.title = element_text(hjust = 0.4),
          axis.title.y = element_blank())
  

ggsave(filename=paste0("RAnalysis/Output/Final_Figures/Host/specific_pathways/amino.DMG.DEG.png"), plot=amino.GO.plot, dpi=300, width=14, height=10, units="in")
ggsave(filename=paste0("RAnalysis/Output/Final_Figures/Host/specific_pathways/amino.DMG.DEG.pdf"), plot=amino.GO.plot, dpi=300, width=14, height=10, units="in")

```

#oxidoreductase activity
```{r}

oxy <- c("oxygen", "oxidative", "oxidoreductase activity")

oxy.DEG <-DEG.complete%>%
      filter(grepl(paste(oxy, collapse="|"), paste(slim_term, term))) 

oxy.DMG <- DMG.complete %>% 
      filter(grepl(paste(oxy, collapse="|"), paste(slim_term, term)))

oxy.dat <- rbind(oxy.DEG,oxy.DMG)

  #plot significantly enriched GO terms by Slim Category faceted by slim term 
  oxy.GO.plot <-  ggplot(oxy.dat, aes(x = type, y = term)) + 
    geom_point(aes(size=numInCat, fill=bh_adjust, colour=bh_adjust)) + 
    scale_size(name="Number of Genes", range=c(5,12), limits=c(1,534))+
    scale_fill_gradient(name = "Adjusted p-value", low="black", high="gray")+
    scale_colour_gradient(name = "Adjusted p-value", low="black", high="gray")+
    facet_grid(slim_term ~ ., scales = "free", space='free', labeller = label_wrap_gen(width = 8, multi_line = TRUE))+
    xlab("Analysis")+
    ggtitle("Oxidoreductase Activity")+
    theme_bw() + 
    theme(panel.border = element_blank(), 
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"),
          strip.background =element_rect(fill="white"),
          strip.text.y = element_text(angle=0, size = 18, face="bold"),
          strip.text.x = element_text(size = 16),
          axis.text.x = element_text(size = 20, angle=45, hjust=1),
          axis.text.y = element_text(size = 16),
          axis.title.x = element_text(size = 22, face="bold"),
          legend.text=element_text(size=16),
          title=element_text(size=22, face="bold"),
          plot.title = element_text(hjust = 0.4),
          axis.title.y = element_blank())
  

ggsave(filename=paste0("RAnalysis/Output/Final_Figures/Host/specific_pathways/oxy.DMG.DEG.png"), plot=oxy.GO.plot, dpi=300, width=18, height=10, units="in")
ggsave(filename=paste0("RAnalysis/Output/Final_Figures/Host/specific_pathways/oxy.DMG.DEG.pdf"), plot=oxy.GO.plot, dpi=300, width=18, height=10, units="in")

```

#protein synthesis
```{r}

protein <- c("protein metabolic process", "protein folding", "protein", "protein modification process", "protein transport")

protein.DEG <-DEG.complete%>%
      filter(grepl(paste(protein, collapse="|"), paste(slim_term, term))) 

protein.DMG <- DMG.complete %>% 
      filter(grepl(paste(protein, collapse="|"), paste(slim_term, term)))

protein.dat <- rbind(protein.DEG,protein.DMG)

  #plot significantly enriched GO terms by Slim Category faceted by slim term 
  protein.GO.plot <-  ggplot(protein.dat, aes(x = type, y = term)) + 
    geom_point(aes(size=numInCat, fill=bh_adjust, colour=bh_adjust)) + 
    scale_size(name="Number of Genes", range=c(5,12), limits=c(1,1500))+
    scale_fill_gradient(name = "Adjusted p-value", low="black", high="gray")+
    scale_colour_gradient(name = "Adjusted p-value", low="black", high="gray")+
    facet_grid(slim_term ~ ., scales = "free", space='free', labeller = label_wrap_gen(width = 30, multi_line = TRUE))+
    xlab("Analysis")+
    ggtitle("Protein Synthesis")+
    theme_bw() + 
    theme(panel.border = element_blank(), 
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"),
          strip.background =element_rect(fill="white"),
          strip.text.y = element_text(angle=0, size = 18, face="bold"),
          strip.text.x = element_text(size = 16),
          axis.text.x = element_text(size = 20, angle=45, hjust=1),
          axis.text.y = element_text(size = 16),
          axis.title.x = element_text(size = 22, face="bold"),
          legend.text=element_text(size=16),
          title=element_text(size=22, face="bold"),
          plot.title = element_text(hjust = 0.4),
          axis.title.y = element_blank())
  

ggsave(filename=paste0("RAnalysis/Output/Final_Figures/Host/specific_pathways/protein.DMG.DEG.png"), plot=protein.GO.plot, dpi=300, width=20, height=24, units="in")
ggsave(filename=paste0("RAnalysis/Output/Final_Figures/Host/specific_pathways/protein.DMG.DEG.pdf"), plot=protein.GO.plot, dpi=300, width=20, height=24, units="in")

```

#glucose metabolism
```{r}

glucose <- c("glucose", "glucagon", "glucokinase", "glucocorticoid", "glucosinolate")

glucose.DEG <-DEG.complete%>%
      filter(grepl(paste(glucose, collapse="|"), paste(slim_term, term))) 

glucose.DMG <- DMG.complete %>% 
      filter(grepl(paste(glucose, collapse="|"), paste(slim_term, term)))

glucose.dat <- rbind(glucose.DEG,glucose.DMG)

  #plot significantly enriched GO terms by Slim Category faceted by slim term 
  glucose.GO.plot <-  ggplot(glucose.dat, aes(x = type, y = term)) + 
    geom_point(aes(size=numInCat, fill=bh_adjust, colour=bh_adjust)) + 
    scale_size(name="Number of Genes", range=c(5,12), limits=c(1,20))+
    scale_fill_gradient(name = "Adjusted p-value", low="black", high="gray")+
    scale_colour_gradient(name = "Adjusted p-value", low="black", high="gray")+
    facet_grid(slim_term ~ ., scales = "free", space='free', labeller = label_wrap_gen(width = 30, multi_line = TRUE))+
    xlab("Analysis")+
    ggtitle("Glucose Metabolism")+
    theme_bw() + 
    theme(panel.border = element_blank(), 
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"),
          strip.background =element_rect(fill="white"),
          strip.text.y = element_text(angle=0, size = 18, face="bold"),
          strip.text.x = element_text(size = 16),
          axis.text.x = element_text(size = 20, angle=45, hjust=1),
          axis.text.y = element_text(size = 16),
          axis.title.x = element_text(size = 22, face="bold"),
          legend.text=element_text(size=16),
          title=element_text(size=22, face="bold"),
          plot.title = element_text(hjust = 0.4),
          axis.title.y = element_blank())
  

ggsave(filename=paste0("RAnalysis/Output/Final_Figures/Host/specific_pathways/glucose.DMG.DEG.png"), plot=glucose.GO.plot, dpi=300, width=18, height=10, units="in")
ggsave(filename=paste0("RAnalysis/Output/Final_Figures/Host/specific_pathways/glucose.DMG.DEG.pdf"), plot=glucose.GO.plot, dpi=300, width=18, height=10, units="in")

```

#glutamate metabolism
```{r}

glutamate <- c("glutamate", "glutathione", "glutamylcyclotransferase", "hydroxymethylglutaryl", "glutathione")

glutamate.DEG <-DEG.complete%>%
      filter(grepl(paste(glutamate, collapse="|"), paste(slim_term, term))) 

glutamate.DMG <- DMG.complete %>% 
      filter(grepl(paste(glutamate, collapse="|"), paste(slim_term, term)))

glutamate.dat <- rbind(glutamate.DEG,glutamate.DMG)

  #plot significantly enriched GO terms by Slim Category faceted by slim term 
  glutamate.GO.plot <-  ggplot(glutamate.dat, aes(x = type, y = term)) + 
    geom_point(aes(size=numInCat, fill=bh_adjust, colour=bh_adjust)) + 
    scale_size(name="Number of Genes", range=c(5,12), limits=c(1,200))+
    scale_fill_gradient(name = "Adjusted p-value", low="black", high="gray")+
    scale_colour_gradient(name = "Adjusted p-value", low="black", high="gray")+
    facet_grid(slim_term ~ ., scales = "free", space='free', labeller = label_wrap_gen(width = 30, multi_line = TRUE))+
    xlab("Analysis")+
    ggtitle("Glutamine Synthetase and Glutamate Synthase Pathway")+
    theme_bw() + 
    theme(panel.border = element_blank(), 
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"),
          strip.background =element_rect(fill="white"),
          strip.text.y = element_text(angle=0, size = 18, face="bold"),
          strip.text.x = element_text(size = 16),
          axis.text.x = element_text(size = 20, angle=45, hjust=1),
          axis.text.y = element_text(size = 16),
          axis.title.x = element_text(size = 22, face="bold"),
          legend.text=element_text(size=16),
          title=element_text(size=22, face="bold"),
          plot.title = element_text(hjust = 0.4),
          axis.title.y = element_blank())
  

ggsave(filename=paste0("RAnalysis/Output/Final_Figures/Host/specific_pathways/glutamate.DMG.DEG.png"), plot=glutamate.GO.plot, dpi=300, width=18, height=10, units="in")
ggsave(filename=paste0("RAnalysis/Output/Final_Figures/Host/specific_pathways/glutamate.DMG.DEG.pdf"), plot=glutamate.GO.plot, dpi=300, width=18, height=10, units="in") 

```

#cellular energy metabolism
```{r}

cell <- c("sugar", "fatty acid", "lipid", "fatty")

cell.DEG <-DEG.complete%>%
      filter(grepl(paste(cell, collapse="|"), paste(slim_term, term))) 

cell.DMG <- DMG.complete %>% 
      filter(grepl(paste(cell, collapse="|"), paste(slim_term, term)))

cell.dat <- rbind(cell.DEG,cell.DMG)

  #plot significantly enriched GO terms by Slim Category faceted by slim term 
  cell.GO.plot <-  ggplot(cell.dat, aes(x = type, y = term)) + 
    geom_point(aes(size=numInCat, fill=bh_adjust, colour=bh_adjust)) + 
    scale_size(name="Number of Genes", range=c(5,12), limits=c(1,40))+
    scale_fill_gradient(name = "Adjusted p-value", low="black", high="gray")+
    scale_colour_gradient(name = "Adjusted p-value", low="black", high="gray")+
    facet_grid(slim_term ~ ., scales = "free", space='free', labeller = label_wrap_gen(width = 40, multi_line = TRUE))+
    xlab("Analysis")+
    ggtitle("Cellular Energy Metabolism") +
    theme_bw() + 
    theme(panel.border = element_blank(), 
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"),
          strip.background =element_rect(fill="white"),
          strip.text.y = element_text(angle=0, size = 18, face="bold"),
          strip.text.x = element_text(size = 16),
          axis.text.x = element_text(size = 20, angle=45, hjust=1),
          axis.text.y = element_text(size = 16),
          axis.title.x = element_text(size = 22, face="bold"),
          legend.text=element_text(size=16),
          title=element_text(size=22, face="bold"),
          plot.title = element_text(hjust = 0.4),
          axis.title.y = element_blank())
  

ggsave(filename=paste0("RAnalysis/Output/Final_Figures/Host/specific_pathways/cell.DMG.DEG.png"), plot=cell.GO.plot, dpi=300, width=18, height=10, units="in")
ggsave(filename=paste0("RAnalysis/Output/Final_Figures/Host/specific_pathways/cell.DMG.DEG.pdf"), plot=cell.GO.plot, dpi=300, width=18, height=10, units="in")

```


#methithionie pathway
```{r}

methionine <- c("methionine", "methionyl", "methionine", "methyl", "SAM", "aden")

methionine.DEG <-DEG.complete%>%
      filter(grepl(paste(methionine, collapse="|"), paste(slim_term, term))) 

methionine.DMG <- DMG.complete %>% 
      filter(grepl(paste(methionine, collapse="|"), paste(slim_term, term)))

methionine.dat <- rbind(methionine.DEG,methionine.DMG)

  #plot significantly enriched GO terms by Slim Category faceted by slim term 
methionine.GO.plot <-  ggplot(methionine.dat, aes(x = type, y = term)) + 
    geom_point(aes(size=numInCat, fill=bh_adjust, colour=bh_adjust)) + 
    scale_size(name="Number of Genes", range=c(5,12), limits=c(1,300))+
    scale_fill_gradient(name = "Adjusted p-value", low="black", high="gray")+
    scale_colour_gradient(name = "Adjusted p-value", low="black", high="gray")+
    facet_grid(slim_term ~ ., scales = "free", space='free', labeller = label_wrap_gen(width = 40, multi_line = TRUE))+
    xlab("Analysis")+
    ggtitle("Methionine Pathway") +
    theme_bw() + 
    theme(panel.border = element_blank(), 
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"),
          strip.background =element_rect(fill="white"),
          strip.text.y = element_text(angle=0, size = 18, face="bold"),
          strip.text.x = element_text(size = 16),
          axis.text.x = element_text(size = 20, angle=45, hjust=1),
          axis.text.y = element_text(size = 16),
          axis.title.x = element_text(size = 22, face="bold"),
          legend.text=element_text(size=16),
          title=element_text(size=22, face="bold"),
          plot.title = element_text(hjust = 0.4),
          axis.title.y = element_blank())
  

ggsave(filename=paste0("RAnalysis/Output/Final_Figures/Host/specific_pathways/methionine.DMG.DEG.png"), plot=methionine.GO.plot, dpi=300, width=18, height=10, units="in")
ggsave(filename=paste0("RAnalysis/Output/Final_Figures/Host/specific_pathways/methionine.DMG.DEG.pdf"), plot=methionine.GO.plot, dpi=300, width=18, height=10, units="in")

```

#use patchwork to combine figures and add labels

```{r}

Figure4 <- glucose.GO.plot + glutamate.GO.plot + oxy.GO.plot + plot_annotation(tag_levels = 'A') & theme(plot.tag = element_text(size = 50))
Figure5 <- nutrient.GO.plot + nitrogen.GO.plot + carb.GO.plot + heat.GO.plot + amino.GO.plot + protein.GO.plot + cell.GO.plot + methionine.GO.plot + plot_annotation(tag_levels = 'A') & theme(plot.tag = element_text(size = 50))
#Figure5 <- heat.GO.plot + amino.GO.plot + protein.GO.plot + cell.GO.plot + plot_annotation(tag_levels = 'A')
 
#ggsave(filename=paste0("RAnalysis/Output/Final_Figures/Host/manuscript/nitrogen.carb_pathways.png"), plot=Figure3, dpi=300, width=40, height=15, units="in")
#ggsave(filename=paste0("RAnalysis/Output/Final_Figures/Host/manuscript/nitrogen.carb_pathways.pdf"), plot=Figure3, dpi=300, width=40, height=15, units="in")

ggsave(filename=paste0("RAnalysis/Output/Final_Figures/Host/manuscript/glut.oxy.pathways.png"), plot=Figure4, dpi=300, width=49, height=20, units="in")
ggsave(filename=paste0("RAnalysis/Output/Final_Figures/Host/manuscript/glut.oxy.pathways.pdf"), plot=Figure4, dpi=300, width=49, height=20, units="in")

ggsave(filename=paste0("RAnalysis/Output/Final_Figures/Host/manuscript/full.fig.png"), plot=Figure5, dpi=300, width=60, height=55, units="in", limitsize = FALSE)
ggsave(filename=paste0("RAnalysis/Output/Final_Figures/Host/manuscript/full.fig.pdf"), plot=Figure5, dpi=300, width=60, height=55, units="in", limitsize = FALSE)
```

