---
title: "GO_Term_Enrichment_Analysis"
author: "daniellembecker"
edited by: "daniellembecker"
date: "9/20/2021"
output: html_document
---

# Molecular Underpinnings Chronic Nutrient Enrichment Project

## RNAseq Gene Ontology (GO) Enrichment Analysis 
## Previous steps include RNAseq workflow in Bioinformatics>RNAseq>RNAseq workflow and Differential Gene Expression statistical analysis in RAnalysis>Scripts>RNAseq_Differential_Gene_Expression to make DMG statistical data sheet
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Load libraries
if (!requireNamespace("DESeq2", quietly = TRUE)) { BiocManager::install("DESeq2") }
library("DESeq2")
library("tidyr")
library("tidyverse")
library("dplyr")
library("pheatmap")
library("RColorBrewer")
if (!requireNamespace("genefilter", quietly = TRUE)) { BiocManager::install("genefilter") }
library("genefilter")
library("ggplot2")
library("gplots")
library("spdep") 
library("patchwork")
#library("aDMGenet") 
library("gridExtra")
if (!requireNamespace("clusterProfiler", quietly = TRUE)) { BiocManager::install("clusterProfiler") }
library("clusterProfiler")
library("DataCombine")
if (!requireNamespace("GSEABase", quietly = TRUE)) { BiocManager::install("GSEABase") }
library("GSEABase")
library("data.table")
if ("rrvgo" %in% rownames(installed.packages()) == 'FALSE') BiocManager::install("rrvgo")
library("rrvgo")
if ("org.Ce.eg.db" %in% rownames(installed.packages()) == 'FALSE') BiocManager::install("org.Ce.eg.db")
library(org.Ce.eg.db)
if ("topGO" %in% rownames(installed.packages()) == 'FALSE') BiocManager::install("topGO")
library(topGO)
if ("Rgraphviz" %in% rownames(installed.packages()) == 'FALSE') BiocManager::install("Rgraphviz")
library(Rgraphviz)
```

# Import the data files 
```{r, echo=FALSE}
# Load go terms from functional annotation file
annot_GO <- read.csv("RAnalysis/Output/annot_GO.terms.host.csv")[-1]

# Remove rows where 'GO.ID' is 'unknown'
annot_GO <- subset(annot_GO, GO.ID != 'unknown')

# edit annot_GO to list GO terms all in one row per gene
annot_GO <- annot_GO %>%
  group_by(gene_id) %>%
  summarise(GO.ID = paste(GO.ID, collapse = ";")) %>%
  ungroup()

```

# Load in DMG results
```{r}
#DMG results
DMG.res <- read.csv("RAnalysis/Output/DMGs_all.csv")[-1]

colnames(DMG.res)[1] <-"gene_id" # make colnames a true column called gene_id
#remove the extra "gene" to match the "Pver_gene_g#" format
DMG.res$gene_id <-  gsub("Pver_gene_novel_gene_","Pver_novel_gene_", DMG.res$gene_id)
DMG.res$gene_id <-  gsub("Pver_gene_split_gene_", "Pver_gene_", DMG.res$gene_id)
#remove extra characters to match annot
DMG.res$gene_id <-  gsub("_gene_g","_g",DMG.res$gene_id) #sub gene_g with _g
DMG.res$gene_id <-  gsub("_gene_","_",DMG.res$gene_id) #remove extra _
DMG.res$gene_id <-  gsub("_split_","_",DMG.res$gene_id) #remove split naming in gene id

# Read in genes of interest file 
genes_of_interest <- read.csv("RAnalysis/Output/gene_list_cleaned.csv")[-1]

# Remove duplicate gene_id entries, keeping the first occurrence
genes_of_interest_unique <- genes_of_interest %>%
  distinct(gene_id, .keep_all = TRUE)

# Filter DEG.res to only include genes of interest
DMG.res_filtered <- DMG.res[DMG.res$gene_id %in% genes_of_interest_unique$gene_id, ]

# Use the new MD.10x file that has now been filtered for SNPs
meth_filt <- readRDS("RAnalysis/Data/filtered_bed_SNP_file.Rdata") #read in info

# Rename gene column to gene_id and select only necessary columns, summarize meth
meth_filt_select <- meth_filt  %>%
  dplyr::select(Sample.ID, meth, gene) %>%
  dplyr::rename(gene_id = gene)

# Reshape the data
meth_wide <- meth_filt_select %>%
  group_by(gene_id, Sample.ID) %>%
  summarize(total_meth = sum(meth, na.rm = TRUE), .groups = "drop") %>%
  pivot_wider(names_from = Sample.ID, values_from = total_meth, values_fill = 0)

#remove the extra "gene" to match the "Pver_gene_g#" format
meth_wide$gene_id <-  gsub("Pver_gene_novel_gene_","Pver_novel_gene_", meth_wide$gene_id)
meth_wide$gene_id <-  gsub("Pver_gene_split_gene_", "Pver_gene_", meth_wide$gene_id)
#remove extra characters to match annot
meth_wide$gene_id <-  gsub("_gene_g","_g", meth_wide$gene_id) #sub gene_g with _g
meth_wide$gene_id <-  gsub("_gene_","_", meth_wide$gene_id) #remove extra _
meth_wide$gene_id <-  gsub("_split_","_", meth_wide$gene_id) #remove split naming in gene id
meth_wide$gene_id <-  gsub("-.*","",meth_wide$gene_id) #remove the extra g number after the dash

# Check if all gene IDs in GFF3 are included in  annot_GO
unmatched_genes <- setdiff(meth_wide$gene_id, annot_GO$gene_id)
# Print the unmatched gene IDs
print(unmatched_genes)

# Only genes that don't match after edited are the unknowns we filtered above
```

Match up genes in gene list file to annotation file
```{r}
names(annot_GO)

gcount_filt2annot = match(meth_wide$gene_id, annot_GO$gene_id) #match genes in gcount_filt to annot_GO

# The following is the number of probes without annotation 
sum(is.na(gcount_filt2annot))

row_nas<-which(is.na(gcount_filt2annot))

#view the genes that do not have a match in the annotation file
missing<-as.data.frame(gcount_filt2annot[row_nas])

```
390 genes in mmethylation data do not have GO annotations

Reduce annotation file to only contain genes detected in our dataset.  
```{r}
filtered_Pverr.annot <- annot_GO[annot_GO$gene_id %in% meth_wide$gene_id, ]
dim(filtered_Pverr.annot)

```

The annotation file now only contains genes that were detected in our dataset that have annotation information. This is 3,520 out of the 3,910 genes in our dataset at 10x coverage. 

Biological Processes topGO

Use topGO for enrichment analysis on Hypermethylated data
```{r}
### Generate vector with names in just the contrast we are analyzing
ID.vector.hyper <- DMG.res_filtered %>%
      filter(methylation_status == "Hypermethylated") %>%
      pull(gene_id)
    
# Get a list of GO Terms for all genes detected in our dataset
GO.terms <- filtered_Pverr.annot %>%
  dplyr::select(gene_id, GO.ID) %>%
  separate_rows(GO.ID, sep = ";")

# Create a gene-to-GO mapping list
geneID2GO <- split(as.character(GO.terms$GO.ID), GO.terms$gene_id)
geneNames <- unique(GO.terms$gene_id)

# Create a named vector for allGenes
allGenes.hyper <- factor(as.integer(geneNames %in% ID.vector.hyper))
names(allGenes.hyper) <- geneNames

# Create a topGOdata object using the custom gene2GO mapping
GOdata.hyper_BP <- new("topGOdata",
              ontology = "BP",
              allGenes = allGenes.hyper,
              nodeSize = 10,
              annot = annFUN.gene2GO, 
              gene2GO = geneID2GO)

GOdata.hyper_BP

```


Run topGO statistical analyses on compiled Hypermethylated data
```{r}
# fisher exact test from topgo manual
fisher.hyper_BP <- runTest(GOdata.hyper_BP, algorithm = "weight", statistic = "fisher")

# Generate a results table
allRes_hyper_BP <- GenTable(GOdata.hyper_BP, 
                      classicFisher = fisher.hyper_BP, 
                      ranksOf = "classicFisher", 
                      topNodes = length(score(fisher.hyper_BP)))

# Convert character columns to numeric
allRes_hyper_BP$classicFisher <- as.numeric(allRes_hyper_BP$classicFisher)

```


Create node figrues of top 5-10 sig terms for Hypermethylated genes
```{r}
# Plot significant nodes
#showSigOfNodes(GOdata.hyper_BP, score(fisher.hyper_BP), firstSigNodes = 5, useInfo = "all")

# Print graph with specific parameters and save to directory
#printGraph(GOdata.hyper_BP, fisher.hyper_BP, firstSigNodes = 5, fn.prefix = "RAnalysis/Output/tGO_graph_Hypermethylated_genesofinterest_BP", useInfo = "def", pdfSW = TRUE)

```


Use topGO for enrichment analysis on Hypermethylated data # NO GENES Hypermethylated, so removed from analyses


Molecular Functions topGO

Use topGO for enrichment analysis on Hypermethylated data
```{r}

# Create a topGOdata object using the custom gene2GO mapping
GOdata.hyper_MF <- new("topGOdata",
              ontology = "MF",
              allGenes = allGenes.hyper,
              nodeSize = 10,
              annot = annFUN.gene2GO, 
              gene2GO = geneID2GO)

GOdata.hyper_MF

```


Run topGO statistical analyses on compiled Hypermethylated data
```{r}
# fisher exact test from topgo manual
fisher.hyper_MF <- runTest(GOdata.hyper_MF, algorithm = "weight", statistic = "fisher")

# Generate a results table
allRes_hyper_MF <- GenTable(GOdata.hyper_MF, 
                      classicFisher = fisher.hyper_MF, 
                      ranksOf = "classicFisher", 
                      topNodes = length(score(fisher.hyper_MF)))

# Convert character columns to numeric
allRes_hyper_MF$classicFisher <- as.numeric(allRes_hyper_MF$classicFisher)

```



Create node figrues of top 5-10 sig terms for Hypermethylated genes
```{r}
# Plot significant nodes
#showSigOfNodes(GOdata.hyper_MF, score(fisher.hyper_MF), firstSigNodes = 5, useInfo = 'all')

# Print graph with specific parameters and save to directory
#printGraph(GOdata.hyper_MF, fisher.hyper_MF, firstSigNodes = 5, fn.prefix = "RAnalysis/Output/tGO_graph_Hypermethylated_genesofinterest_MF", useInfo = "def", pdfSW = TRUE)

```




Cellular Components topGO

Use topGO for enrichment analysis on Hypermethylated data
```{r}

# Create a topGOdata object using the custom gene2GO mapping
GOdata.hyper_CC <- new("topGOdata",
              ontology = "CC",
              allGenes = allGenes.hyper,
              nodeSize = 10,
              annot = annFUN.gene2GO, 
              gene2GO = geneID2GO)

GOdata.hyper_CC
```


Run topGO statistical analyses on compiled Hypermethylated data
```{r}
# fisher exact test from topgo manual
fisher.hyper_CC <- runTest(GOdata.hyper_CC, algorithm = "weight", statistic = "fisher")

# Generate a results table
allRes_hyper_CC <- GenTable(GOdata.hyper_CC, 
                      classicFisher = fisher.hyper_CC, 
                      ranksOf = "classicFisher", 
                      topNodes = length(score(fisher.hyper_CC)))

# Convert character columns to numeric
allRes_hyper_CC$classicFisher <- as.numeric(allRes_hyper_CC$classicFisher)

```


Create node figrues of top 5-10 sig terms for Hypermethylated genes
```{r}
# Plot significant nodes
#showSigOfNodes(GOdata.hyper_CC, score(fisher.hyper_CC), firstSigNodes = 5, useInfo = 'all')

# Print graph with specific parameters and save to directory
#printGraph(GOdata.hyper_CC, fisher.hyper_CC, firstSigNodes = 5, fn.prefix = "RAnalysis/Output/tGO_graph_Hypermethylated_CC", useInfo = "def", pdfSW = TRUE)

```

Biological Processes topGO

Use topGO for enrichment analysis on Hypomethylated data
```{r}
### Generate vector with names in just the contrast we are analyzing
ID.vector.hypo <- DMG.res_filtered %>%
      filter(methylation_status == "Hypomethylated") %>%
      pull(gene_id)
    
# Get a list of GO Terms for all genes detected in our dataset
GO.terms <- filtered_Pverr.annot %>%
  dplyr::select(gene_id, GO.ID) %>%
  separate_rows(GO.ID, sep = ";")

# Create a gene-to-GO mapping list
geneID2GO <- split(as.character(GO.terms$GO.ID), GO.terms$gene_id)
geneNames <- unique(GO.terms$gene_id)

# Create a named vector for allGenes
allGenes.hypo <- factor(as.integer(geneNames %in% ID.vector.hypo))
names(allGenes.hypo) <- geneNames

# Create a topGOdata object using the custom gene2GO mapping
GOdata.hypo_BP <- new("topGOdata",
              ontology = "BP",
              allGenes = allGenes.hypo,
              nodeSize = 10,
              annot = annFUN.gene2GO, 
              gene2GO = geneID2GO)

GOdata.hypo_BP

```


Run topGO statistical analyses on compiled Hypomethylated data
```{r}
# fisher exact test from topgo manual
fisher.hypo_BP <- runTest(GOdata.hypo_BP, algorithm = "weight", statistic = "fisher")

# Generate a results table
allRes_hypo_BP <- GenTable(GOdata.hypo_BP, 
                      classicFisher = fisher.hypo_BP, 
                      ranksOf = "classicFisher", 
                      topNodes = length(score(fisher.hypo_BP)))

# Convert character columns to numeric
allRes_hypo_BP$classicFisher <- as.numeric(allRes_hypo_BP$classicFisher)

```


Create node figrues of top 5-10 sig terms for Hypomethylated genes
```{r}
# Plot significant nodes
#showSigOfNodes(GOdata.hypo_BP, score(fisher.hypo_BP), firstSigNodes = 5, useInfo = "all")

# Print graph with specific parameters and save to directory
#printGraph(GOdata.hypo_BP, fisher.hypo_BP, firstSigNodes = 5, fn.prefix = "RAnalysis/Output/tGO_graph_Hypomethylated_genesofinterest_BP", useInfo = "def", pdfSW = TRUE)

```


Use topGO for enrichment analysis on hypomethylated data # NO GENES HYPOMETHYLATED, so removed from analyses


Molecular Functions topGO

Use topGO for enrichment analysis on Hypomethylated data
```{r}

# Create a topGOdata object using the custom gene2GO mapping
GOdata.hypo_MF <- new("topGOdata",
              ontology = "MF",
              allGenes = allGenes.hypo,
              nodeSize = 10,
              annot = annFUN.gene2GO, 
              gene2GO = geneID2GO)

GOdata.hypo_MF

```


Run topGO statistical analyses on compiled Hypomethylated data
```{r}
# fisher exact test from topgo manual
fisher.hypo_MF <- runTest(GOdata.hypo_MF, algorithm = "weight", statistic = "fisher")

# Generate a results table
allRes_hypo_MF <- GenTable(GOdata.hypo_MF, 
                      classicFisher = fisher.hypo_MF, 
                      ranksOf = "classicFisher", 
                      topNodes = length(score(fisher.hypo_MF)))

# Convert character columns to numeric
allRes_hypo_MF$classicFisher <- as.numeric(allRes_hypo_MF$classicFisher)

```



Create node figrues of top 5-10 sig terms for Hypomethylated genes
```{r}
# Plot significant nodes
#showSigOfNodes(GOdata.hypo_MF, score(fisher.hypo_MF), firstSigNodes = 5, useInfo = 'all')

# Print graph with specific parameters and save to directory
#printGraph(GOdata.hypo_MF, fisher.hypo_MF, firstSigNodes = 5, fn.prefix = "RAnalysis/Output/tGO_graph_Hypomethylated_genesofinterest_MF", useInfo = "def", pdfSW = TRUE)

```




Cellular Components topGO

Use topGO for enrichment analysis on Hypomethylated data
```{r}

# Create a topGOdata object using the custom gene2GO mapping
GOdata.hypo_CC <- new("topGOdata",
              ontology = "CC",
              allGenes = allGenes.hypo,
              nodeSize = 10,
              annot = annFUN.gene2GO, 
              gene2GO = geneID2GO)

GOdata.hypo_CC


```


Run topGO statistical analyses on compiled Hypomethylated data
```{r}
# fisher exact test from topgo manual
fisher.hypo_CC <- runTest(GOdata.hypo_CC, algorithm = "weight", statistic = "fisher")

# Generate a results table
allRes_hypo_CC <- GenTable(GOdata.hypo_CC, 
                      classicFisher = fisher.hypo_CC, 
                      ranksOf = "classicFisher", 
                      topNodes = length(score(fisher.hypo_CC)))

# Convert character columns to numeric
allRes_hypo_CC$classicFisher <- as.numeric(allRes_hypo_CC$classicFisher)

```


Create node figrues of top 5-10 sig terms for Hypomethylated genes
```{r}
# Plot significant nodes
#showSigOfNodes(GOdata.hypo_CC, score(fisher.hypo_CC), firstSigNodes = 5, useInfo = 'all')

# Print graph with specific parameters and save to directory
#printGraph(GOdata.hypo_CC, fisher.hypo_CC, firstSigNodes = 5, fn.prefix = "RAnalysis/Output/tGO_graph_Hypomethylated_CC", useInfo = "def", pdfSW = TRUE)

```




Print Tables
```{r}

# Generate a results table
allRes_hyper_BP_10 <- GenTable(GOdata.hyper_BP, 
                      classicFisher = fisher.hyper_BP, 
                      ranksOf = "classicFisher", 
                      topNodes = 10)
allRes_hyper_BP_10$methylation_status <- "Hypermethylated"

# Generate a results table
allRes_hypo_BP_10 <- GenTable(GOdata.hypo_BP, 
                      classicFisher = fisher.hypo_BP, 
                      ranksOf = "classicFisher", 
                      topNodes = 10)
allRes_hypo_BP_10$methylation_status <- "Hypomethylated"

# Generate a results table
allRes_hyper_MF_10 <- GenTable(GOdata.hyper_MF, 
                      classicFisher = fisher.hyper_MF, 
                      ranksOf = "classicFisher", 
                      topNodes = 10)
allRes_hyper_MF_10$methylation_status <- "Hypermethylated"

# Generate a results table
allRes_hypo_MF_10 <- GenTable(GOdata.hypo_MF, 
                      classicFisher = fisher.hypo_MF, 
                      ranksOf = "classicFisher", 
                      topNodes = 10)
allRes_hypo_MF_10$methylation_status <- "Hypomethylated"

# Combine top results for hypo and hyperreagulated BP
Enrichment_Res_BP_10 <- rbind(allRes_hyper_BP_10,allRes_hypo_BP_10)

# Create publishable table
Enrichment_Res_BP_10_table <- Enrichment_Res_BP_10 %>%
  dplyr::select(-Expected, -methylation_status)  # Remove the 'expected' and 'methylation_status" column

# Create flextable object
flex_enrich <- flextable(Enrichment_Res_BP_10_table)

flex_enrich <- flex_enrich %>%
  set_header_labels(
    GO.ID = "GO ID",
    Term = "Term",
    Annotated = "Annotated",
    Significant = "Significant",
    classicFisher = "Fisher p-value") %>%
  colformat_double(digits = 4) %>%
  bold(j = "classicFisher", i = ~ as.numeric(classicFisher) < 0.05) %>%  # Bold values < 0.055, including scientific notation
  add_header_lines(values = "Top 10 Enriched GO Terms (Biological Process)") %>%
  autofit() %>%
  theme_box()

# Save flextable as PNG
png_file <- "RAnalysis/Output/DMG_top10_enrich_BP_table_genesofinterest.png"
save_as_image(flex_enrich, path = png_file)

# Combine top results for hypo and hyperreagulated MF
Enrichment_Res_MF_10 <- rbind(allRes_hyper_MF_10,allRes_hypo_MF_10)

# Create publishable table
Enrichment_Res_MF_10_table <- Enrichment_Res_MF_10 %>%
  dplyr::select(-Expected, -methylation_status)  # Remove the 'expected' and 'methylation_status" column

# Create flextable object
flex_enrich <- flextable(Enrichment_Res_MF_10_table)

# Customize the flextable
flex_enrich <- flex_enrich %>%
  set_header_labels(
    GO.ID = "GO ID",
    Term = "Term",
    Annotated = "Annotated",
    Significant = "Significant",
    classicFisher = "Fisher p-value") %>%
  colformat_double(digits = 4) %>%
  bold(j = "classicFisher", i = ~ as.numeric(classicFisher) < 0.05) %>%  # Bold values < 0.055, including scientific notation
  add_header_lines(values = "Top 10 Enriched GO Terms (Molecular Functions)") %>%
  autofit() %>%
  theme_box()

# Save flextable as PNG
png_file <- "RAnalysis/Output/DMG_top10_enrich_MF_table_genesofinterest.png"
save_as_image(flex_enrich, path = png_file)
```