---
title: "RNA_seq_DESeq2"
author: "echille"
edited by: "daniellembecker"
date: "4/12/2021"
output: html_document
---

# Molecular Underpinnings Chronic Nutrient Enrichment Project

## RNAseq Differential Expression Analysis 
## Follow RNAseq workflow in Bioinformatics>RNAseq>RNAseq workflow before statistical analysis to make gene counts and transcript counts matrices

### Set up workspace

Load libraries
```{r, message=FALSE, warning=FALSE}
library("genefilter")
library("DESeq2")
library("factoextra")
library("NbClust")
library("ComplexHeatmap")
library("tidyverse")
library("RColorBrewer")
library("ggplot2")
library("goseq")
library("gridExtra")
library("VennDiagram")
library("here")
```

Import the data files 
```{r}
#treatment information
treatmentinfo <- read.csv("../Data/RNA-seq/metadata.RNAseq.csv", header = TRUE, sep = ",")
rownames(treatmentinfo) <- treatmentinfo$sample_id
str(treatmentinfo)
head(treatmentinfo)

#gene count matrix
gcount <- as.data.frame(read.csv("../Data/RNA-seq/Poc_gene_count_matrix.csv", row.names="gene_id"))
dim(gcount)
head(gcount)

#remove extra characters from multiple column header names for sample ids, skip first column labeled gene counts so specify 1:ncol
for ( col in 1:ncol(gcount)){
  colnames(gcount)[col] <-  sub("_R1_001.fastq.gz.sam.sorted.bam.merge.gtf", "", colnames(gcount)[col])
}

#subset the gcount matrix and reorder it so it matches the treatment info
gcount <- gcount[ ,treatmentinfo$sample_id]
head(gcount)

# Make sure treatment info ='s the rows of count data
all(rownames(treatmentinfo) %in% colnames(gcount)) # must come out TRUE

```


##### Pre-filter gene counts
## Pre-filtering our dataset to reduce the memory size dataframe, increase the speed of the transformation and testing functions, and improve quality of statistical analysis by removing low-coverage counts. Removed counts could represent outliers in the data and removing these improves sensitivity of statistical tests. We will  filter out low coverage samples. Here, we will keep P=87.5% percent of the samples have counts over A=5, allowing only 1 of 8 samples to have a value less than 5 per gene.

```{r}
#Set filter values for PoverA, P=87.5% percent of the samples have counts over A=5. We chose this value allowing only 1 of 8 samples to have a value less than 5 per gene. 
filt <- filterfun(pOverA(0.875,5))

#create filter for the counts data
gfilt <- genefilter(gcount, filt)

#identify genes to keep by count filter
gkeep <- gcount[gfilt,]

#identify gene lists
gn.keep <- rownames(gkeep)

#gene count data filtered in PoverA, P percent of the samples have counts over A
gcount_filt <- as.data.frame(gcount[which(rownames(gcount) %in% gn.keep),])
head(gcount_filt)
dim(gcount_filt)


```

##Merge the treatment columns into a new column , group. Set group as a factor.
```{r}
treatmentinfo$treatment <- factor(treatmentinfo$treatment, levels = c("enriched","control"))
```

#Create a DESeqDataSet design from gene count matrix and labels. Here we set the design to look at treatment to test for any differences in gene expression across timepoints attributed to treatment.
#not using DESeq due to random factors, but can still use this to visualize clusters
```{r}
#Set DESeq2 design
gdds <- DESeqDataSetFromMatrix(countData = gcount_filt,
                              colData = treatmentinfo,
                              design = ~treatment)
```

#### Visualize gene count data

We're looking to see if the samples of the same treatments cluster

##### Log-transform the count data
First we are going to log-transform the data using a variance stabilizing transforamtion (vst). This is only for visualization purposes. Essentially, this is roughly similar to putting the data on the log2 scale. It will deal with the sampling variability of low counts by calculating within-group variability (if blind=FALSE). Importantly, it does not use the design to remove variation in the data, and so can be used to examine if there may be any variability do to technical factors such as extraction batch effects.

To do this we first need to calculate the size factors of our samples. This is a rough estimate of how many reads each sample contains compared to the others. In order to use VST (the faster log2 transforming process) to log-transform our data, the size factors need to be less than 4. Otherwise, there could be artefacts in our results.
```{r}
SF.gdds <- estimateSizeFactors(gdds) #estimate size factors to determine if we can use vst  to transform our data. Size factors should be less than four to use vst
print(sizeFactors(SF.gdds)) #View size factors
```

Our size factors are all less than 4, so we can use VST!
```{r}
gvst <- vst(gdds, blind=FALSE) #apply a variance stabilizing transformation to minimize effects of small counts and normalize wrt library size
```

#### Principal component plot of samples
```{r}
gPCAdata <- plotPCA(gvst, intgroup = c("treatment"), returnData=TRUE)
percentVar <- round(100*attr(gPCAdata, "percentVar")) #plot PCA of samples with all data
ggplot(gPCAdata, aes(PC1, PC2, color=treatment)) + 
  geom_point(size=3) +
  xlab(paste0("PC1: ",percentVar[1],"% variance")) +
  ylab(paste0("PC2: ",percentVar[2],"% variance")) +
  scale_color_manual(values = c(control="slateblue", enriched="indianred3")) +
  coord_fixed() +
  theme_bw() + #Set background color
  theme(panel.border = element_blank(), # Set border
                     #panel.grid.major = element_blank(), #Set major gridlines
                     #panel.grid.minor = element_blank(), #Set minor gridlines
                     axis.line = element_line(colour = "black"), #Set axes color
        plot.background=element_blank()) + #Set the plot background
  theme(legend.position = ("none")) #set title attributes
```



#### Differential Gene Expression Analysis
# For data with random effects 




