---
title: "dendogram"
author: "daniellembecker"
date: "12/01/23"
output: html_document
---

# Molecular Underpinnings Chronic Nutrient Enrichment Project

## Dendogram of clusters of GO and KEGG terms, enrichment output

#load libraries
```{r}
library(ggplot2)
library(dendextend)
library(plyr)
library(circlize)
library(cowplot)
library(gridExtra)
library(data.table)
library(stringdist)
library(tidytree)
library(png)
library(grid)
library(ggpubr)
library(viridis)
library(treeio)
library(ggtree)
library(patchwork)
library(stringr)
library(readr)
```

#input data
```{r}
f_gene_groups=c('RAnalysis/Data/WGBS/dendogram/dendogram_DMG_MF.csv') 
```

#output directory + base name for the output graph
```{r}
outidx=c('RAnalysis/Output/Dendograms/DMG_clusters.graphs')

f_gene_groups_sep=","
```
 

#names of columns in the input csv
```{r}
gene_groups_ids='genes'
gene_groups_direction='percentDEInCat' #calculated as (number of DEGs in each GO/total number of genes in GO term)*100 _this is the percentage of differentially expressed 

#genes in each term. It will be used to set the color in the heatmap
gene_groups_term='term'
gene_groups_experiment='experiment'
gene_groups_size='set_size'


```

# functions:

```{r}
dist_funct=function(v,func){
  n=length(v)
  m=matrix(data=NA,nrow=n,ncol=n)
  for(j in 1:n){
    if(j<n){
      for(k in (j+1):n){
        dist = func(v[j],v[k])
        m[j,k]= dist
        m[k,j]= dist
      }
    }
  }
  return(m)
}

dist_overlap=function(x,y){
  x=strsplit(x,';')[[1]]
  y=strsplit(y,';')[[1]]
  z=length(intersect(x,y))
  return(1-(z/min(length(x),length(y))))
}

pairwise_dists=function(ids,dist_mat_){
  dist_mat=as.matrix(dist_mat_)
  ids_hash=list()
  rownames0=rownames(dist_mat_)
  for(j in 1:length(rownames0)){ids_hash[[rownames0[j]]]=j}
  dists=c()
  n=length(ids)
  for(j in 1:n){
    if(j<n){
      for(k in (j+1):n){
        dists=c(dists,dist_mat[ids_hash[[ids[j]]],ids_hash[[ids[k]]]])
      }
    }
  }
  return(dists)
}



colors1=c('black','blue','brown','burlywood4','cadetblue','chartreuse4','chocolate3','chocolate4','coral3',
          'blue4','blueviolet','brown1','darkgoldenrod4','deepskyblue4','firebrick4','darkslategrey','purple','red',
          'green','gray',rainbow(n=100))  #colors for the dendrogram clusters
max1=5  #you can try different numbers and choose the best clustering 
```


##### run a for loop for each term to make connections for dendogram
```{r}
for(i in 1:length(f_gene_groups)){
  #geneGroups1 = read.csv(f_gene_groups[i],sep=f_gene_groups_sep,stringsAsFactors=F)
  geneGroups1 = fread(f_gene_groups[i],sep=f_gene_groups_sep,stringsAsFactors=F)
  cols=c(gene_groups_ids,gene_groups_direction,gene_groups_term,gene_groups_experiment, gene_groups_size)
  geneGroups1 = geneGroups1[,..cols,with=F]
  names(geneGroups1) = c('ids','score','term','experiment', 'size')
  geneGroups2 = geneGroups1[,.(nonredudnatIDs = paste(unique(strsplit(gsub("\\s+","",paste(ids,collapse=';'),perl=T)[[1]],";")[[1]]),collapse=";")),by=term]
  geneGroups3 = as.data.frame(geneGroups2)
  d = dist_funct(geneGroups3[,'nonredudnatIDs'],dist_overlap)
  colnames(d)=geneGroups3[,'term']
  rownames(d)=geneGroups3[,'term']
  hc = hclust(as.dist(d),method="ward.D2") #hclust_method)
  hcd = as.dendrogram(hc)
  hc[['tip.label']] = hc[['labels']] }
  
  
  cuts0=1:max1
  cutr1=list()
  cutr_clusters1=list()
  for(u in 1:length(cuts0)){
    cat(u,"\n")
    cutr1[[u]] = cutree(tree=hc,k=cuts0[u])
    cutr_freqs = as.data.frame(table(cutr1[[u]]))
    cutr_clusters1[[u]]=data.frame(term_name=names(cutr1[[u]]),cluster=cutr1[[u]])
    cutr_clusters1[[u]]=cutr_clusters1[[u]][order(cutr_clusters1[[u]]$cluster),]
  }
  
  idx0 = as.integer(max1)
  cutr=cutr1[[idx0]]
  cuts=cuts0[idx0]
  cutr_clusters=cutr_clusters1[[idx0]]
  cutr_tightClusters=cutr_clusters1[[length(cutr_clusters1)]]
  names(cutr_tightClusters) = c('term','cluster')
  
 # Uppercase the first letter of row names in the data frame
cutr_tightClusters$label_upper <- tools::toTitleCase(rownames(cutr_tightClusters))

# make clusters on tree and organize
gg1 = ggtree(hc, layout = 'rectangular') %<+% cutr_tightClusters + #edits layout of tree
  geom_tippoint(aes(color = as.factor(cluster)), shape = "triangle", size = 2) + #adds triangles and adjusts size
  scale_colour_manual(values = colors1) + #distinguishes colors of clusters with data
  geom_tiplab(aes(label = label_upper), offset = 0.1, size = 4) + #references the label_upper column to name tree nodes
  xlim(-3, 3) #sets x limites on plot
  
  pdf(paste0(outidx[i],'_tree_MF.pdf'))
  print(gg1)
  dev.off()
  
  scores1 = as.data.frame(dcast(geneGroups1[,.(term,score,experiment)], term ~ experiment,value.var='score',mean),stringsAsFactors=F)
  scores2 = scores1[,c(1:ncol(scores1))]
  scores2 = as.data.frame(scores2)
  scores2 <- select(scores2, -c("term"))
  rownames(scores2) = scores1[,1]
  gg2 = gheatmap(gg1, scores2, offset=1.75, width=0.1, font.size= 5) 
  gg2 = gg2 + viridis::scale_fill_viridis(na.value=NA, option = "inferno")
  pdf(paste0(outidx[i],'_tree_scores_MF.pdf'),width=12, height = 8)
  print(gg2)
  dev.off() 


```


#make dot plot seperate from dendorgram
```{r}
#calculate means for set size and percentDEIncat for each term
# Calculate mean score for each term
mean_scores <- geneGroups1 %>%
  group_by(term) %>%
  summarise(mean_score = mean(score))

# Calculate sum of set size for each term
sum_set_sizes <- geneGroups1 %>%
  group_by(term) %>%
  summarise(sum_set_size = sum(size))

#combine data frames by term
dotplot <- left_join(mean_scores, sum_set_sizes)

#read in dotplot data
clusters <- read_csv("RAnalysis/Data/WGBS/dendogram/DMG_dotplot_MF.csv", col_names = TRUE)


gg3 = ggplot (clusters, aes(x = comparison, y = term, size = set_size)) + 
            geom_count (aes(color = percentDEInCat)) + 
            scale_color_viridis(na.value=NA, option = "inferno") +
            scale_size(limits = c(4, 5600)) +  # Set the limits here
            labs(color = "%DMGs", size = "Number of Genes") +
            theme(panel.background = element_blank())

print(gg3)

ggsave("RAnalysis/Output/Dendograms/clusters_dotplot_DMG_MF.png",
       dpi = 300, 
       width = 8,
       height = 5)

```


#make figure one for manuscript
```{r}
#load png image of DMG PCA

#load images of BP DEG dendogram and MF DEG dendogram
dendo.BP <- readPNG("RAnalysis/Output/Dendograms/DMG_biological_processes.png")
dendo.BP2 <- rasterGrob(dendo.BP, interpolate=TRUE)

#load images of MF DEG dendogram
dendo.MF <- readPNG("RAnalysis/Output/Dendograms/DMG_molecular_functions.png")
dendo.MF2 <- rasterGrob(dendo.MF, interpolate=TRUE)

Figure4 <- plot_grid(dendo.BP2, dendo.MF2, labels = c('A', 'B'), label_size = 14, nrow = 2)

ggsave(filename=paste0("RAnalysis/Output/Final_Figures/Host/manuscript/dendogram_DMG.png"), plot=Figure4, dpi=300,width=4, height=5, units="in", limitsize = FALSE)
ggsave(filename=paste0("RAnalysis/Output/Final_Figures/Host/manuscript/dendogram_DMG.pdf"), plot=Figure4, dpi=300, width=13, height=5, units="in", limitsize = FALSE)

```







